### 2015-05-08 article.tex
initial commit

\documentclass[5p]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\modulolinenumbers[5]

\journal{Journal of Neural Networks}

\bibliographystyle{bst/elsarticle-num}

\begin{document}

\begin{frontmatter}

\title{V4 Neural Network Model for Shape-based Feature Extraction and Object Discrimination}

\author{Zheng Dong}
\ead{dongzheng@fudan.edu.cn}

\author{Hui Wei\corref{cor:1}}
\cortext[cor:1]{Corresponding author}
\ead{weihui@fudan.edu.cn}

\address{Laboratory of Cognitive Modeling and Algorithms, %
School of Computer Science, Fudan University, Shanghai, China}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

Understanding the content of images has always been a difficult task in image analysis
due to the well known semantic gap \cite{smeulders2000}.
A high degree of abstraction is required to transform low-level pixel data into high-level knowledge of the images.
Computer vision researchers attempt to bridge the gap between pixel data and image content with image features.
Various methods have been proposed to extract features from pixel representation of images.
However, humans and primates outperform any existing computer vision system with respect to almost any measure.
How visual cortex recognizes objects is a critical question for neuroscience.
Visual stimuli captured by the eye are transformed into neural impulses in the retina and further processed by 
the visual cortex for high level cognitive tasks.
The visual cortex plays an important role in filling the gap
between visual stimuli and the implied semantic information.
Neural impulses travel through the visual pathway \cite{ettlinger1990}
in the visual cortex and finally contribute to a unified percept of the object of interest.

Primate brains possess two distinct visual pathways \cite{ettlinger1990,lehky2007}.
The dorsal pathway is involved with processing the object's spatial location relevant to the viewer. 
The ventral pathway is involved with object discrimination and recognition.
In this paper, we concentrate on the latter, the ventral pathway, and visual area V4 in particular.
V4 lies in the middle of the ventral pathway (\figurename~\ref{fig:1}).
Lower levels of the pathway (visual areas V1 and V2) extract preliminary visual features 
including edges and orientations \cite{hubel1962,hubel1965}.
Higher levels of the pathway (inferior temporal cortex) exhibit selectivity to complex objects
like faces and body parts \cite{bruce1981,bell2009}.
As an intermediate stage, V4 plays a crucial role in transforming low-level orientation signals 
into complex object representations.

\begin{figure}
\centerline{\includegraphics[width=0.8\linewidth]{images/fig-1.png}} 
\caption{V4 is an intermediate stage in visual recognition.}
\label{fig:1}
\end{figure}

In this paper, we propose a multi-layer neural network model
which imitates the neural mechanism of visual area V4 in the ventral visual pathway.
Different stages of the neural pathway in the visual cortex
are explicitly mapped to different layers of the neural network model.
In this feed-forward network, features are extracted hierarchically from input images. 
High-level local image features emerge in the output layer of the network. 
The features encode local structures of object shapes in a way which closely resembles the mechanism of the cortex 
and therefore achieve a good balance between the discrimination and generalization abilities.
The features exhibit competitive performance when applied to object recognition tasks.
Moreover, this multilayer neural network shows a successful attempt 
to integrate different stages of visual processing in the brain. 
It provides an extensible framework so that future findings in cortex mechanisms 
can be adopted and applied to computer vision tasks.

We briefly introduce some related work on traditional image features
and the biological background of our model in section~\ref{sec:2}.
We focus on the shape selectivity of area V4 and some previous models of V4 are also discussed in this section.
In section~\ref{sec:3}, we give the detailed description of our proposed model.
It is capable of locating salient points 
and generating discriminative local shape features around each salient point.
In section~\ref{sec:4}, empirical evaluation shows that
the model demonstrates similar selectivity compared with V4 neurons.
and compared with other models in section~\ref{sec:exp}. 
The conclusion is summarized in section~\ref{sec:conclusion}.

\section{Related work}\label{sec:2}

\subsection{Traditional image features}

Image features encode the image structure in a spatial neighborhood around each of a set of interest points.
Extracting features is an important stage in the process of understanding the content of an image.
These features can be low-level features, such as edges \cite{canny1986} 
and corners \cite{harris1988,smith1997},
or high-level features, which are usually more distinctive and descriptive.

Traditional image features mostly fall into three categories \cite{mikolajczyk2005}: 
distribution-based descriptors, spatial-frequency descriptors and differential-based descriptors.
Distribution-based descriptors use histograms 
to represent characteristics of local image regions.
SIFT descriptor \cite{lowe1999} (including its variations) 
and shape context \cite{belongie2002} are typical examples of such descriptors.
SIFT descriptor uses histograms 
of oriented image gradients over rectangular grids.
Shape context implements similar idea 
by computing histograms of edge pixels in a log-polar grid.
Spatial-frequency descriptors and differential-based descriptors 
use filters and derivatives to capture the invariant features. 
Examples of these categories are 
complex filters \cite{schaffalitzky2002} 
and Baumberg's affine invariant descriptor \cite{baumberg2000}.
Complex filters are a family of filters 
derived by adding complex multipliers to a Gaussian filter.
Baumberg's descriptor is generated 
by applying an elliptical Gaussian filter 
to the second order moment 
of the scale-space derivative \cite{lindeberg1993} of images.

These traditional image features utilize flat models which compute features directly from the image
via statistical or algebraic techniques. 
The flat approach does not conform to the hierarchical mechanism \cite{ungerleider2011} 
employed by neural systems such as the primate brain.

\subsection{Neural mechanism of vision}

At lower levels of the ventral visual pathway (V1 and V2), neurons respond to features of local orientation. 
These neurons fall into two categories, 
simple cells and complex cells \cite{hubel1962,hubel1965,martinez2003}.
The receptive fields of simple cells 
can be understood as linear filters modeled as Gabor functions \cite{gabor1946}. 
They respond primarily to oriented edges and gratings.
Complex cells differ from simple cells in that a stimulus is
effective wherever it is placed in the receptive field, provided
that the orientation is appropriate \cite{hubel1962,hubel1965}.

Neurons in V4 and IT cortex encode high-level information.
The exact nature of the encoded information remains to be determined.
The response pattern of V4 neurons shows 
a strong bias towards convex contour features \cite{pasupathy1999,pasupathy2001,pasupathy2002}. 
V4 neurons are selective for orientation and curvature of boundary conformation.
Research also suggests that simple shapes are represented 
by patterns of activity across populations of V4 neurons \cite{pasupathy2002}.

IT cortex is significant in visual recognition and
discrimination \cite{desimone1980,bell2009,ungerleider2011,bruce1981}.
It depends exclusively on inputs received from
connections originating in V1 \cite{desimone1980}.
Recently, object representations in the temporal cortex of monkeys and humans are revealed 
by functional magnetic resonance imaging \cite{bell2009}.
Discrete regions selective for animate and inanimate stimuli were found in both species.
These regions could be further subdivided into regions selective for individual categories.

\subsection{Shape Selectivity of V4 Neurons}

Neurobiological studies on V4 have not produced a unified model of its function or circuitry.
V4 neurons are known to be selective for color, shape, depth and even motion \cite{roe2012}.
In this paper, we focus on the V4 selectivity for shapes. 
Early experiments examined the selectivity of cells in V4 
with classical stimuli including bars and sinusoidal gratings~(\figurename~\ref{fig:2a}) \cite{desimone1987}.
Similar to earlier processing stages, some V4 neurons are tuned for orientation 
and spatial frequency of edges and linear sinusoidal gratings.
However, the majority of V4 neurons are sensitive to more complex shape properties. 
Later experiments demonstrated that V4 neurons display a clear bias in their responses 
in favor of non-Cartesian gratings~(\figurename~\ref{fig:2b}) and they show a significant degree of invariance 
in their selectivity across changes in stimulus position \cite{gallant1996}.
More recent experiments showed that V4 neurons can be strongly selective for curvature
of contours and angular position of acute curvatures \cite{pasupathy1999,pasupathy2001}. 
\figurename~\ref{fig:2c} shows a response map of a V4 neuron
(reproduced from \cite{pasupathy2001}).
The white shapes in \figurename~\ref{fig:2c} is presented in the receptive field of this V4 neuron
with equally dark background. 
The gray scale in the response map indicates the strength of the neuronal response.
Darker background indicates that the neuronal response is stronger.
This neuron is selectively tuned for acute convex border at the bottom right.
The curvature and the angular position are both necessary conditions of the activation of this neuron.
Neither rounded protrusions nor sharp curvatures towards directions other than the bottom left
activate the neuron.

\begin{figure}
\centering
\subfloat[]{\includegraphics[width=0.18\linewidth]{images/fig-2a.png}\label{fig:2a}}\hfil
\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2b.png}\label{fig:2b}}\hfil
\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2c.png}\label{fig:2c}}
\caption{Shapes to examine V4 selectivity.
(a) Classical gratings.
(b) Non-Cartesian gratings. V4 neurons prefer non-Cartesian gratings rather than classical gratings.
(c) Response map of a V4 neuron which responds to a sharp convex curvature at the bottom left.
Darker background indicates a stronger response.}
\label{fig:2}
\end{figure}

\subsection{Previous models of V4}

Several models have been proposed for V4 neurons.
We briefly introduce the SRF model \cite{david2006} and the HMAX model \cite{riesenhuber1999,cadieu2007}.

The spectral receptive field (SRF) \cite{david2006}
describes properties of V4 receptive field in terms of the orientation and spatial frequency spectrum.
The model is based on the fact that V4 neurons have large orientation and spatial frequency bandwidth.
They respond selectively to stimuli such as contour conformations and non-Cartesian gratings,
which generally consist of multiple orientations and spatial frequencies.
The spectral model is also invariant to small changes in stimulus position 
and thus explains the invariance property of V4 response patterns.
The model is powerful in describing the shape selectivity of V4 neurons.
However, it does not explain the emergence of the selectivity.
It does not either provide the neural computing process to achieve such spectral receptive field.

The HMAX model was first proposed in \cite{riesenhuber1999}
as a generic model for object recognition in the visual cortex.
It models the visual cortex into a hierarchical architecture 
consisting of cascaded linear filters and non-linear maximum pooling operations.
It was then adopted as a model for V4 shape selectivity and invariance \cite{cadieu2007}.
The training of the model is an NP-complete problem. 
The authors used a greedy algorithm to obtain approximated solutions
but they did not provide any biological evidence for the algorithm.

The proposed model of V4 in this paper overcomes the limitation of the previous models.
In the next section, we demonstrate that its architecture and function is analogous 
to those of the visual cortex.
We also provide efficient training method for our model.

\section{Multi-layer neural network model}\label{sec:3}

The information processing in the visual cortex follows a hierarchical scheme.
Our model employs a similar hierarchical structure.
Different cortex areas along the visual pathway are explicitly adopted in our model. 
\figurename~\ref{fig:3} shows the architecture of our model.

\subsection{Innovation in our model}

The application of neural science in computer vision has been very limited.
The most widely used DoG filters and Gabor filters 
are models of ganglion cells in the retina and simple cells in the primary visual cortex, 
which correspond to early stages of visual processing in the brain.

Biologically inspired models \cite{serre2007, hong2011},
proposed recently have exhibited competitive performance in traditional computer vision tasks.
However, these models are still limited to the utilization 
of the structure of the primary visual cortex. 
Most of these models originate from the HMAX \cite{riesenhuber1999} model 
which duplicates the structure of simple cells and complex cells
in the primary visual cortex for high-level processing.

In our model, we attempt to map each stage in the ventral pathway of visual processing
to a single layer of computational units.
These units are carefully designed so that they conform to the neural mechanism.
We focus on the neural mechanism of V4 with particular concern
because the shape features extracted in V4 are significant in the process of object discrimination.

\begin{figure}
\centerline{\includegraphics[width=\linewidth]{images/fig-3.pdf}} 
\caption{Multilayer neural network model of V4.}
\label{fig:3}
\end{figure}

\subsection{Layer of simple cells}

According to the hierarchy of the ventral visual pathway,
area V4 receives input from the lower levels including area V1 and V2.
These areas have been well studied since 1960s 
by Hubel, Wiesel \cite{hubel1962,hubel1965} and succeeding researchers.

Neurons in V1 and V2 respond to local orientations.
They fall into two categories, simple cells and complex cells.
Simple cells respond primarily to oriented edges and gratings.
Complex cells have larger receptive fields.
A stimulus is effective wherever it is placed in the complex receptive field, 
provided that the orientation is appropriate \cite{hubel1962}.

In our model, the layer of simple cells covers the whole dimension of input images.
For each location, several simple cells with different preferred orientations 
detect edges of different orientations.
Simple cells are modeled as Gabor functions \cite{gabor1946},
\begin{equation}\label{equ:gabor}
g_{\theta}(x,y;\lambda,\sigma_s)
=\exp \left(-\frac{x'^2+y'^2}{2\sigma_s^2}\right)
\cos \left(2\pi\frac{x'}{\lambda}+\psi\right),
\end{equation}
where $x'=x\cos\theta+y\sin\theta$, $y'=-x\sin\theta+y\cos\theta$.
In the equation, $\lambda$ is the wavelength of the sinusoidal factor, 
$\psi$ is the phase offset, 
$\theta$ represents the preferred orientation and 
$\sigma_s$ approximates the radius of the receptive fields.

\begin{figure}
\centering
\subfloat[Phase offset $\psi=\pi/2$]{
  \includegraphics[width=0.4\linewidth]{images/fig-4a.pdf}
  \label{fig:4a}}
\subfloat[Phase offset $\psi=0$]{
  \includegraphics[width=0.4\linewidth]{images/fig-4b.pdf}
  \label{fig:4b}}
\caption{Gabor functions with different phases.}
\label{fig:4}
\end{figure}

It is notable that the phase offset $\psi$ affects the pattern of the receptive fields.
\figurename~\ref{fig:4}~shows two Gabor functions with different phase offsets.
The sinusoidal factors of the two Gabor functions 
are odd function and even function respectively.
The odd Gabor function resembles a gradient operator
while the even Gabor function resembles a Laplacian operator.
In a computational point of view, 
the Laplacian operator calculates a higher order derivative 
than the gradient operator.
The gradient operator suppresses the smooth area in images.
Moreover, the Laplacian operator is able 
to suppress area that varies smoothly, 
which is also neglected by human eyes generally.
Therefore, in our model, the even Gabor function 
with the phase offset $\psi=0$ is employed.

The layer of simple cells operates on raw image input. 
The output of simple cells with preferred orientation $\theta$
and scale $\sigma_s$ is the following convolution passed through a transfer function $\phi$,
\begin{equation}\label{equ:gabor}
S_{\theta,\sigma_s}(x,y)=\phi(I\otimes g_{\theta,\sigma_s}),
\end{equation}
where $I$ is an image and 
\begin{equation}\label{equ:gabor}
\phi(u)=\left\{\begin{array}{ll}
u & \text{if } u>0,\\
0 & \text{otherwise.}
\end{array}\right.
\end{equation}

\subsection{Layer of complex cells}

Complex cells are modeled as Gaussian filters which take inputs from simple cells,
\begin{equation}
f(x,y;\sigma_c)=\frac{1}{2\pi\sigma_c^2}\exp\left(-\frac{x^2+y^2}{2\sigma_c^2}\right),
\end{equation}
where $\sigma_c$ represents the radius of the complex receptive fields.
The filters are discretized into matrices and normalized before applied to digital images.
The radius of complex receptive field is approximately twice as long as that of a simple cell.
We take this size because qualitatively the complex receptive fields are usually two to five times wider 
than the bars of gratings that stimulate them most effectively \cite{movshon1978}.
In this paper, we have $\sigma_c=2\sigma_s$.

Given $I(x,y)$ as some input image, the output of complex cells
with the preferred orientation $\theta$ is defined as the following convolution.
\begin{equation}
C_{\theta}=|S_{\theta}|\otimes f.
\label{equ:complex}
\end{equation}
We use absolute values of simple cell outputs 
because we do not distinguish between dark edges and bright edges.
Generally speaking, simple cells modeled as in equation (\ref{equ:gabor}) 
with phase offset $\psi=0$ response best to stimuli of bright stripes in dark background, i.e., bright edges.
Such cells are thus called ON-center \cite{hubel1962} cells in the neural system.
Dark edges result in a negative response (strong inhibition) of ON-center cells.
In the neural system, OFF-center cells with inversed receptive fields response to dark edges.
In our model, ON-center and OFF-center cells are not distinguished.
In other words, the layer of complex cells in our model is sensitive to bright edges, dark edges 
and boundaries between bright and dark areas.

\subsection{Saliency filters in V4 layer}

V4 is an area of attentional modulation \cite{roe2012}.
Visual attention involves selecting an interested region or selecting specific object features.
Visual attention in V4 is influenced by both top-down feedback from higher levels in the visual pathway
and bottom-up input from lower levels.
We focus on the bottom-up influence.
In a bottom-up process, V4 evaluates the saliency of the input from lower levels
and focuses its attention automatically on the salient features.

We use entropy to measure the saliency of images \cite{kadir2001}.
In \cite{kadir2001}, it is assumed that salient regions have high complexity (and correspondingly high entropy).The entropy of features is used as a scale invariant measure.
The salient region is selected at entropy peaks over scales.
To avoid erroneously selecting noise or texture as salient regions,
a measure of self-similarity is employed.
Self-similar regions are filtered out.

In this paper, we use entropy in a different approach.
V4 neurons encode fragments of object contour \cite{pasupathy2001,pasupathy2002}.
They are selective for simple structures such as convex or concave curvature.
Therefore, in our assumption, 
well ordered structures with low complexity are preferred in cognitive activities.
We filter out regions with high complexity (or entropy).

The entropy is calculated according to the output of complex cells.
Given a point $(x,y)$, in the neighborhood of $(x,y)$,
a complex cell with preferred orientation $\theta$ has output value $C_{\theta}(x,y)$,
We suppose that the probability of a complex cell being activated
is proportional to the output value.
The probability is thus defined as:
\begin{equation}
P(\theta)=\frac{1}{\sum_{\theta_i} C_{\theta_i}(x,y)}\cdot C_{\theta}(x,y).
\end{equation}
The entropy of the complex cell activity in this neighborhood
is 
\begin{equation}
E=-\sum_{\theta} P(\theta) \log P(\theta).
\end{equation}

\figurename~\ref{fig:5} shows four image patches.
For each image patch, the output values of complex cells with different preferred orientations
(from $0^\circ$ to $180^\circ$) are plotted in a bar chart.
The charts show the distribution of complex cell activities.
The entropy calculated accordingly indicates that 
patches composed of simple structures have non-uniform distributions and thus low entropy values.
Therefore, we filter out regions with high entropy.

\begin{figure}
\centerline{\includegraphics[width=0.8\linewidth]{images/fig-5.png}} 
\caption{Entropy of image patches.
Bar charts in the second row show the output values of complex cells
with different preferred orientations (from $0^\circ$ to $180^\circ$).}
\label{fig:5}
\end{figure}

In addition to entropy, 
local competition also plays a role in attentional selection \cite{desimone1995}.
With limited neural resources, only strong and competitive neuronal signals get transmitted and processed.
In our model, the saliency layer finds local maximums of complex cell output
and filters out those points with high entropy values or low activities (or output values).
The algorithm is listed in Algorithm \ref{alg:1}.

\begin{algorithm}
  \caption{Saliency filter}
  \label{alg:1}
  \begin{algorithmic}[1]
    \Procedure{FindSalientPoint}{image $I$, scale $\sigma_c$, threshold $t_A,t_E$}
      \For{each orientation $\theta$}
        \State $C_{\theta}\leftarrow\phi(I\otimes g_{\theta})\otimes f_{\sigma_c}$
        \Comment{complex cell output}
      \EndFor
      \For{each point $(x,y)$ in image $I$}
        \State $C(x,y)\leftarrow\max_{\theta}C_{\theta}(x,y)$
        \State $E(x,y)\leftarrow$ entropy at point $(x,y)$
      \EndFor
      \State Divide image $I$ into patches of size $\sigma_c\times\sigma_c$
      \For{each patch $p$}
        \State $(\hat{x},\hat{y})\leftarrow\operatorname{argmax}_{(x,y)\in p}C(x,y)$
        \If{$C(\hat{x},\hat{y})>t_A$ and $E(\hat{x},\hat{y})<t_E$}
          \State Mark $(\hat{x},\hat{y})$ as a salient point
        \EndIf
      \EndFor
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\subsection{RBM encoders in V4 layer}

With saliency filters described in the previous subsection,
we are able to focus on a limited number of salient points.
The V4 computation units in our model encode the shape in the neighborhood of each salient point.
The encoding is achieved with Restricted Boltzmann Machine (RBM).

RBM can learn a probability distribution over its set of inputs.
It has been found efficient in training deep neural network \cite{hinton2006}.
The encoder layer in our model is part of a deep network.
Therefore, we use the RBM as a training model for this encoder layer.
The RBM is trained to encode shape features in the neighborhood of a salient point.
\figurename~\ref{fig:6} shows such an RBM encoder.

Let $(x,y)$ denote a salient point.
For each preferred orientation, 
we select the complex cell of which the receptive field is centered at point $(x,y)$,
and its eight neighbors.
\figurename~\ref{fig:6} demonstrates an example of three preferred orientations.
In this case, $9\times3$ complex cells are selected.
The number of selected neurons depends on the number of orientations we choose.
The selected complex cells are then used as the input layer (or visible layer) of the RBM.
The representation of the shape in the neighborhood of $(x,y)$
is formed in the output layer (or hidden layer) of the RBM.

\begin{figure}
\centerline{\includegraphics[width=0.99\linewidth]{images/fig-6.pdf}} 
\caption{RBM encoder for local shape feature.}
\label{fig:6}
\end{figure}

RBM can be trained efficiently with a contrastive divergence learning algorithm \cite{hinton2002}.
Let $w_{ij}$ be the weight of the connection from the $i$-th visible unit to the $j$-th hidden unit.
In each learning iteration, the change in the weight is given by
\begin{equation}
\Delta w_{ij}=\epsilon\left(\langle v_i h_j\rangle_\text{data}-\langle v_i h_j\rangle_\text{recon}\right),
\end{equation}
where $\epsilon$ is a learning rate, 
$\langle v_i h_j\rangle_\text{data}$ is the product of two units when the visible layer is given real data,
and $\langle v_i h_j\rangle_\text{recon}$ is the product of two units 
when the visible layer is given reconstructed data.

We use weight-decay \cite{hinton2010}
to reduce overfitting by adding an L2 penalty term, $\frac{1}{2}\lambda w^2$.
The weight change is then given by
\begin{equation}
\Delta w_{ij}=\epsilon\left(\langle v_i h_j\rangle_\text{data}-\langle v_i h_j\rangle_\text{recon}\right)
-\epsilon\lambda w_{ij},
\end{equation}
where $\lambda$ is a weight-cost coefficient.
We followed \cite{hinton2010} for the choice of the coefficient $\lambda$ and the learning rate $\epsilon$
in our experiments.

The output vector of the hidden layer forms a representation of local shape features.
It can be used directly as input for classification tasks.
We can also use the RBM connection weight matrix to initialize a multilayer neural network
for supervised back-propagation training.

\subsection{Model Parameters}

The size of the receptive field increases along the hierarchy of the visual system.
Lower levels have relatively small receptive fields while higher levels have larger receptive fields.
The size of receptive field in our model follows the same scheme.
The relationship is shown in Table \ref{tab:1}.
V4 neurons receive afferent connections from $3\times3$ complex cells
that have partially overlapped receptive fields.
Therefore the radius of the V4 receptive field is not 3 times as long as that of the complex cell.

\begin{table}[h]
\caption{Size of receptive field}
\centering
\begin{tabular}{rr}
\toprule
Category & Radius of receptive field \\
\midrule
Simple cells & $\sigma_s$ \\
Complex cells & $\sigma_c = 2\sigma_s$ \\
V4 RBM encoders & $2.5\sigma_c$\\
\bottomrule
\end{tabular}
\label{tab:1}
\end{table}

We used two schemes for the number of different preferred orientations of simple cells and complex cells.
For black and white images of shapes and handwritten digits in our experiment,
we used 4 orientations (from $0^\circ$ to $135^\circ$ in steps of $45^\circ$).
For gray scale images, we used 18 orientations (from $0^\circ$ to $170^\circ$ in steps of $10^\circ$)
in order to preserve more information.

The output of complex cells was normalized to the $[0,1]$ interval
to serve as the input of RMB encoders.
The saliency filters took a threshold $t_E=2.8$ for the entropy
and a threshold $t_A=0.4$ for complex cell output value (or complex cell activity).
These values were roughly the median values of natural images.

\section{Shape-based feature extraction}\label{sec:4}

In this section, we demonstrate a series of experiments.
In the experiments, our proposed model extracts local shape features from images.
Compared with V4 neurons, it shows similar selectivity for different features.

\subsection{Perceptron over Complex Cells}

We have reviewed the shape selectivity of V4 neurons in section \ref{sec:2}.
In the following experiment, we examined that the output of complex cells provides sufficient information
for the emergence of neuronal response pattern of V4.
The experiment involved only the first two layers (i.e., simple cells and complex cells) in our model.
A perceptron was configured to take complex cells as input.
It was then trained to distinguish between two shapes shown in \figurename~\ref{fig:7a}.
The shapes were also used to examine the selectivity of V4 in \cite{pasupathy2001}.
The difference between the two shapes is that one has a sharp projection towards the top right.

\begin{figure}
\centering
\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{images/fig-7a.png}\label{fig:7a}}\hfil
\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{images/fig-7b.png}\label{fig:7b}}
\caption{Response map and weight matrices of a perceptron that distinguishes two shapes.
(a) Responses of the perceptron over 18 samples.
The perceptron prefers the shape in the left half of the samples.
It is insensitive to stimulus position.
(b) The input weight of the perceptron. 
Each block shows the weight of connections from complex cells of a certain orientation.
Complex cells of 18 different orientations provide input for this perceptron.}
\label{fig:7}
\end{figure}

Since V4 neurons show a certain degree of invariance in their selectivity across changes in stimulus position,
we moved the shapes randomly within the receptive field of the perceptron 
to generate samples for training and testing (\figurename~\ref{fig:7a} shows several samples).
The samples were then used as input of the layers of simple cells and complex cells.
The output of complex cells were passed to the perceptron.
We had complex cells with different preferred orientations 
(from $0^\circ$ to $170^\circ$ in steps of $10^\circ$) and over different positions in the receptive field.
Therefore, the input of the perceptron consisted of 18 matrices, 
each corresponding to the output of complex cells with a certain orientation.
The input weight of the perceptron was thus also 18 matrices.

The trained perceptron exhibited a strong bias towards the shape with convex curvature towards the top right.
It also showed a significant degree of invariance to the stimulus position.
It is obvious that the selectivity is not formed from certain excitatory sub-regions
or inhibitory sub-regions of the receptive field which was found of simple cells \cite{hubel1962,pasupathy2001}.
The selectivity of the perceptron tallies with the selectivity of actual V4 neurons.
The response map is shown in \figurename~\ref{fig:7a}.
Darker background colors indicate stronger responses.
The input weight of this perceptron is shown in \figurename~\ref{fig:7b}.
Each block shows the weight from the complex cells of a certain preferred orientation.
Red color denotes positive weight while blue color denotes negative weight.

This experiment demonstrates that complex cells provide sufficient information
for V4 neurons to show selectivity observed in neurobiological experiments.

\subsection{Shape Selectivity}

In the following experiment, we trained our model to learn shapes.
The training process involved the layers of simple cells, complex cells and RBM encoders.
The training samples were collected from \cite{gallant1996,pasupathy1999,pasupathy2001}, 
including 4 categories of stimuli, i.e.,
sinusoidal gratings, non-Cartesian gratings, segmented curves, and closed shapes.
The simple cells in our model took these sample images as input.
The images were processed by simple cells and complex cells.
Since the sample images were fitted into V4 receptive fields in neurobiological experiments,
we adjusted the scale of the images to fit the size of a single V4 receptive field in our model too.
Therefore the saliency filters were not necessary in this experiment.
The output of complex cells were passed directly on to an RBM encoder.

We used an RBM with 256 hidden units (RBM output neurons) in this experiment.
In order to assess the selectivity of these units,
we assigned each unit a selectivity index over a category of stimuli.
The selectivity index is defined as the ratio of the maximal response to the average response 
over the stimuli of a certain category.
Given an RBM output unit $h$ and its output value $h_i$ over a category of $N$ stimuli for $i=1,2,\dots,N$.
The selectivity index $\mathcal{S}$ of $h$ is given by
\begin{equation}
\mathcal{S}(h)=\frac{\max\{h_i|i=1,2,\dots,N\}}{\sum_{i=1}^N h_i/N}.
\end{equation}
In the following statistics, we assumed that a neuron has significant selectivity over a category of stimuli
if the selectivity index is greater than 3.5.
Among the 256 units, 171 units exhibited significant selectivity for different stimuli.
102 units showed selectivity for segmented curves (\figurename~\ref{fig:8}).
55 units showed strong bias towards non-Cartesian gratings (\figurename~\ref{fig:9})
while 50 units showed selectivity for classical sinusoidal gratings.
68 units exhibited selectivity for closed shapes (\figurename~\ref{fig:10}).
The shape selectivity of our model is listed in \tablename~\ref{tab:2}.
The categories of selectivity are not mutually disjoint.
A single unit may possess more than one kind of selectivity simultaneously.

\begin{table}[h]
\caption{Shape selectivity of our model}
\centering
\begin{tabular}{rrr}
\toprule
Selectivity & Number of Units & Percentage \\
\midrule
Sinusoidal gratings & 50 & 19.5\% \\
Non-Cartesian gratings & 55 & 21.5\% \\
Segmented curves & 102 & 39.8\% \\
Closed shapes & 68 & 26.6\% \\
\bottomrule
\end{tabular}
\label{tab:2}
\end{table}

\figurename~\ref{fig:8a} shows the sample patterns of segmented curves.
\figurename~\ref{fig:8b} to \figurename~\ref{fig:8d} show the response maps of 3 RBM output neurons.
Darker background indicates stronger response.
The selectivity indexes is shown at the top.
It is shown that the three neurons exhibited strong bias towards curves rather than straight lines.
The neurons were tuned for the orientation of the projection of the segmented curves.
This is compliant with the selectivity of V4 neurons described in \cite{pasupathy1999}.

\figurename~\ref{fig:9} shows the response maps of two RBM output neurons
which are selective for non-Cartesian gratings.
Both of the two neurons were not sensitive to sinusoidal gratings.
The one in \figurename~\ref{fig:9a} was selective for a pair of circular sectors.
The one in \figurename~\ref{fig:9b} preferred non-Cartesian gratings of helix shapes.

\figurename~\ref{fig:10} shows the sample patterns of closed shapes and the response map of two neurons.
The two neurons were selectively tuned for two kinds of shapes.
They were sensitive to the angular position of the shapes.

\begin{figure}
\centering
\subfloat[Patterns]{\includegraphics[width=0.24\linewidth]{images/fig-8a.png}\label{fig:8a}}\hfil
\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8b.png}\label{fig:8b}}\hfil
\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8c.png}\label{fig:8c}}\hfil
\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8d.png}\label{fig:8d}}
\caption{RBM output neurons responding to convex curves towards certain directions.
(a) Sample patterns of segmented curves.
(b) Response map of a neuron tuned for curves projecting upwards.
(c) Response map of a neuron tuned for curves projecting downwards.
(d) Response map of a neuron tuned for curves projecting towards the top left.}
\label{fig:8}
\end{figure}

\begin{figure}
\centering
\subfloat[Response map of a neuron selective for a pair of circular sectors (selectivity index = 3.92).]%
{\includegraphics[width=0.9\linewidth]{images/fig-9a.png}\label{fig:9a}}\\
\subfloat[Response map of a neuron selective for helix shapes (Selectivity index = 6.37).]%
{\includegraphics[width=0.9\linewidth]{images/fig-9b.png}\label{fig:9b}}
\caption{Response maps of RBM output neurons tuned for non-Cartesian gratings.}
\label{fig:9}
\end{figure}

\begin{figure}
\centering
\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{images/fig-10a.png}\label{fig:10a}}\hfil
\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10b.png}\label{fig:10b}}\hfil
\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10c.png}\label{fig:10c}}
\caption{RBM output neurons responding to closed shapes.}
\label{fig:10}
\end{figure}

This experiment demonstrates that our model can learn to distinguish the stimuli
which V4 neurons are selectively tuned for.

\subsection{Feature detection}

In the above experiments, it has been demonstrated that a unit of the RBM hidden layer 
can encode the shape feature in the receptive field of a single V4 computation unit.
When we deal with larger images, especially natural images,
a full scan over the whole image is not computationally economic.
It also complicates succeeding processing for high-level tasks such as object recognition and scene understanding.
We have to find some regions of interest 
and focus the computation units over a limited number of interesting regions.
The visual neural system takes a similar approach.
As we have reviewed in previous sections, V4 is closely related with selective visual attention \cite{roe2012}.
Research shows that V4 receptive field shrinks and shifts towards saccade target \cite{hamker2006}.
We simulated such visual attention with the saliency filters in our model.
In the following experiment,
we evaluated the saliency filters with feature detection experiment.

\begin{figure}
\centering
\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{images/fig-11a.png}\label{fig:11a}}\hfil
\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11c.png}\label{fig:11b}}\hfil
\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11e.png}\label{fig:11c}}\\
\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{images/fig-11b.png}\label{fig:11d}}\hfil
\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11d.png}\label{fig:11e}}\hfil
\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11f.png}\label{fig:11f}}\\
\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11g.pdf}\label{fig:11g}}\hfil
\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11h.pdf}\label{fig:11h}}
\caption{Feature detection. (a) and (d) show the original images.
(b) and (e) are the complex cell activity. 
Red color indicates strong activity and blue color indicates inhibition.
(c) and (f) are the entropy map.
Red color indicates high entropy value and blue color indicates low entropy value.
(g) and (h) are the selected salient points
at the scale of 4, 8, and 16 in terms of simple cell radius.}
\label{fig:11}
\end{figure}

\figurename~\ref{fig:11} shows the process of feature detection with the saliency filters in our model.
\figurename~\ref{fig:11b} and \figurename~\ref{fig:11d} show the activity map of complex cells.
The activities are high in the region of the trees behind the car (\figurename~\ref{fig:11a})
and the handwriting on the white board (\figurename~\ref{fig:11c}).
Saliency filters filter out these regions because these regions have a comparatively high entropy.
These areas consist of a large amount of disordered edges
which result in a near uniform distribution of complex cell activity over preferred orientations
and thus high entropy values.
The saliency filters can find salient points at each given scale.
\figurename~\ref{fig:11g} and \figurename~\ref{fig:11h} show the salient points
at the scale of 4, 8, and 16 in terms of the simple cell radius
(The size of the other units can be inferred with the relationship
shown in Table \ref{tab:1}).

\section{Object discrimination}

In this section, we demonstrate the performance of the features 
generated by our model in feature matching and object recognition tasks.

We do not use typical criteria for evaluating local descriptors \cite{mikolajczyk2005}
for the reason that the typical criteria focus on
matching the same physical point in different viewing conditions,
at which humans or other intelligent species
are not significantly better than computer systems.
The human vision excels in image understanding tasks
such as recognizing a new instance of some category
with only few prototypes for learning.
Our model conforms to the neural mechanism of vision and
thus exhibits similar learning ability in the experiments.

\subsection{Feature matching}

Traditional local features excel in matching features
corresponding to the same physical point on an object
seen from different viewpoints \cite{mikolajczyk2005}.
These features are usually very selective for a target object
but lack invariance with respect to different objects from a target category.
However, the generalization ability within object categories
is important for image understanding tasks such as object recognition.

We provide a qualitative evaluation of our features
by matching features between different images.
The SIFT features are used for comparison.
The VLFeat open-source library \cite{vedaldi2008}
were utilized for the SIFT implementation 
and the feature matching algorithm.

\begin{figure}[h]
\centering
  \subfloat[Matching between the same face]{
    \includegraphics[width=0.8\linewidth]{images/fig-12a.pdf}
    \label{fig:12a}}\\
  \subfloat[Matching between different faces]{
    \includegraphics[width=0.8\linewidth]{images/fig-12b.pdf}
    \label{fig:12b}}
\caption{Matching face images.}
\label{fig:12}
\end{figure}

\figurename~\ref{fig:12} shows matchings of features in face images.
As shown in \figurename~\ref{fig:12a},
SIFT features match perfectly between face images of the same person.
The neural features generated by our model achieve competitive distinctiveness.
However, \figurename~\ref{fig:12b} shows that
matching of SIFT features between face images of different people
is quite disordered.
SIFT features excel in matching highly distinctive features under image transformations 
but lack the generalization ability 
to capture variations in objects appearance of the same category.
By contrast, the neural features show robust performance 
in matching features from different faces.
The result suggests that our model exhibits 
both competitive discrimination ability
and better generalization ability compared with SIFT features.

\subsection{IT layer}

The exact nature of the high-level visual features 
encoded by IT neurons remains to be determined.
Yet research shows that V4 neurons encode shapes 
by population responses \cite{Pasupathy2002}.
It is likely that an IT neuron collects 
the responses from a group of afferent V4 neurons 
and synthesizes a unified representation of shapes.
Thus in our model,
an IT unit combines the inputs 
from several V4 neurons into a local image feature.
The receptive fields of the connected V4 neurons 
scatter around the center of the IT receptive field.
Though there is no anatomical evidence showing that 
the spatial distribution of the connected V4 receptive fields 
is necessarily restricted to any fixed patterns,
in our model, we choose afferent V4 neurons
from a group of rectangular grid points.
Each IT neuron has 9 afferent connections from the V4 layer.
The 9 connected neurons are located in a $3\times 3$ grid.
\figurename~\ref{fig:it} demonstrates an IT receptive field 
and the receptive fields of the connected V4 neurons.
The demonstration is shown over
the output of complex cells given the image of Lena as input.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{images/fig3c.eps}
\caption{Demonstration of IT neuron in our model. 
IT neuron combines V4 inputs in a grid layout rotated to the strongest orientation.}
\label{fig:it}
\end{figure}

The gird is rotated to the strongest orientation 
of the central V4 input 
in order to achieve the invariance of feature orientations.
We were inspired by the SIFT \cite{Lowe1999} descriptor.
The rotation is also demonstrated in \figurename~\ref{fig:it}.
The grid is rotated to the orientation of the edge of the hat
appearing in the center of the grid.

The output of IT neurons are thus defined as a concatenation of the V4 input vectors.
\begin{equation}
\mathbf{\vec{F}}(x,y) = \mathbf{\vec{V}}(\hat{x}_1,\hat{y}_1)\|
\mathbf{\vec{V}}(\hat{x}_2,\hat{y}_2)\|
\mathbf{\vec{V}}(\hat{x}_3,\hat{y}_3)\|{}  \cdots\|
\mathbf{\vec{V}}(\hat{x}_n,\hat{y}_n),
\end{equation}
where $(\hat{x}_i,\hat{y}_i)^{\top} = 
\left(\begin{array}{ccc}
\cos\hat\theta & -\sin\hat\theta & x \\
\sin\hat\theta & \cos\hat\theta & y
\end{array}\right)\cdot(x_i,y_i,1)^{\top}$ for $i=1,2,3,\dots,n$ and 
$\hat\theta = \arg\!\max_\theta C_\theta(x,y)$.
In this equation, $\mathbf{\vec{F}}$ represents the image feature generated by IT neurons, 
$\hat\theta$ represents the strongest orientation of the central V4 input, 
$(\hat{x}_i,\hat{y}_i)$ is a group of rotated grid points
indicating the centers of receptive fields of afferent V4 neurons, 
and $(x_i,y_i)$ is a group of grid points.
In our model, $(x_i,y_i)$ is defined as the $3\times 3$ grid points fitted 
in a rectangular region.

\subsection{Object recognition}\label{sec:exp:objectrecognition}

We evaluate the performance of the proposed model
in object recognition in clutter for which 
the target object in both the training and test samples
appears at variable scales and positions 
within an unsegmented image.

The bags of key-points approach \cite{Csurka2004} are employed.
Key-points are detected with the salient region detector 
proposed by Kadir and Brady \cite{Kadir2001}.
Features at the key-points are extracted with our model
and passed to a classifier.
We experimented with the SVM classifier
and evaluated the classifier over the Caltech 101 dataset \cite{Fei-Fei2004}.
The parameters of our model are summarized in \tablename~\ref{tab:param}.

\begin{table}
\caption{Model Parameters}
\label{tab:param}
\begin{center}
\begin{tabular}{ll}
\hline
{\bf Description} &{\bf Value}\\ 
\hline
Radius of salient region & $R$ \\
Grid size of IT unit ($|x_{i}-x_{i+1}|$) & $|x_{i}-x_{i+1}|=R$ \\
Radius of complex receptive field ($\sigma_c$) & $\sigma_c=R/3$ \\
Radius of simple receptive filed ($\sigma_s$) & $\sigma_s=\sigma_c/2$ \\
\hline
\end{tabular}
\end{center}
\end{table}

In addition to the parameters summarized in \tablename~\ref{tab:param},
the sampling density of different preferred orientations $\theta_k$
($k=1,2,3,\dots$) as in equation (\ref{equ:v4}) is also a parameter in our model.
A density of $n$ means that $\theta_k$ is taken from the range
$\{0^{\circ}, 180^{\circ}/n, 2\times 180^{\circ}/n, \dots, (n-1)\times 180^{\circ}/n\}$.
We investigate the selection of sampling density of preferred orientations.
30 training samples are randomly chosen from each category.
For each training image, 50 features are randomly chosen and passed to the classifier.
The results reported are averaged over 8 independent runs.

\begin{figure}[h]
\centering
  \subfloat[Sampling density of orientations]{
    \includegraphics[width=0.8\linewidth]{images/fig5a.eps}
    \label{fig:result:density}}\\
  \subfloat[Number of features per image]{
    \includegraphics[width=0.8\linewidth]{images/fig5b.eps}
    \label{fig:result:nfeature}}
\caption{Performance with different parameters.}
\label{fig:result}
\end{figure}

As shown in \figurename~\ref{fig:result:density},
the selection of sampling density does not exhibit obvious impact
on the performance.
With a small density, we can obtain 
both relatively low-dimensional feature vectors
and satisfactory performance.
Thus in the succeeding experiments, a sampling density of $3$ is used.

To investigate the contribution of the number of features 
on performance,
we vary the number of features chosen from each training image.
We choose 20 training samples from each category.
The results reported are also averaged over 8 independent runs.
\figurename~\ref{fig:result:nfeature} shows that 
with 50 features per image the result approaches the best performance.
The number of required features can be notably reduced by introducing 
a clustering step in the bag of words approach \cite{Csurka2004}.
However, Serre et al. \cite{Serre2007} reported that clustering could
be sensitive to the choice of parameters and initializations,
leading poorer results.

The influence of the number of training samples 
on the performance of our model is also studied.
For each category, we randomly select training sets of size
5,10,15,20,25 and 30.
\figurename~\ref{fig:result:ntrain} shows the performance as
a function of the number of training images.
The results reported are averaged over all categories
with 50 features extracted from each training image.
The performance is compared with other published results
\cite{Bosch2007,Bosch2007b,Zhang2007,Lazebnik2006,Wang2006,Grauman2005}.

\begin{figure}[h]
\centering
  \subfloat[Comparison with other methods]{
    \includegraphics[width=0.8\linewidth]{images/fig5c.eps}
    \label{fig:result:ntrain}}\\
  \subfloat[Comparison with other features]{
    \includegraphics[width=0.8\linewidth]{images/fig5d.eps}
    \label{fig:result:compare}}
\caption{Comparison of performance.}
\label{fig:result2}
\end{figure}

\figurename~\ref{fig:result:compare} shows
the performance of our model over the 101 categories
compared with SIFT features \cite{Lowe1999},
PHOW features \cite{Lazebnik2006},
and biologically inspired HMAX model \cite{Serre2007}.
The results are obtained with 30 training images randomly selected 
from each category.
SIFT features and neural features generated with our model
are passed to an SVM classifier.
The images passed to the HMAX model are resized 
to a uniform height of 140 pixels as recommended by \cite{Serre2007}.
Our model exhibits competitive performance 
compared with other models.
As shown in \figurename~\ref{fig:result:compare},
the SIFT performance points over the 101 categories
are almost completely to the left of the equivalent line.
The PHOW performance points spread through 
a wider range from $0.5$ to $1$ indicating 
the lack of robustness compared with our model.
Compared with the biologically inspired HMAX model,
our performance is also better for more than half
of the 101 categories.

\section{Conclusion}\label{sec:conclusion}

In this paper, we make an attempt 
to construct a multi-layer neural network model 
that explicitly maps different layers 
to different stages in the visual pathway of object recognition. 
We demonstrate in experiments that local image features 
generated by this neural model show competitive performance 
in object discrimination. 
More over, this model provides a framework 
to take into account higher-level stages in visual processing, 
such as the area V4 and the IT cortex, 
and emulates object recognition mechanisms in the cortex.

There are at least two directions that could be
followed to further improve the model.
First, feedback connections should be
adopted for the reason that feedback connections
occur widely in the neural systems in the forms 
of both intra-layer and inter-layer feedbacks.
Second, computational units that better 
conform to the neural response patterns should
be well evaluated and adopted 
to replace the current coarse implementation 
of the V4 and IT layers.

\section*{References}

\bibliography{ref}

\end{document}

### 2015-05-19 article.tex
before conclusion

@@
 
-\subsection{IT layer}
-
+\subsection{Object recognition layer}
+
+As we have reviewed in section 2,
+the IT cortex is where object recognition is finally achieved in the brain.
 The exact nature of the high-level visual features 
 encoded by IT neurons remains to be determined.
-Yet research shows that V4 neurons encode shapes 
-by population responses \cite{Pasupathy2002}.
-It is likely that an IT neuron collects 
-the responses from a group of afferent V4 neurons 
+Yet research shows that V4 neurons encode shapes by population responses \cite{pasupathy2002}.
+It is likely that an IT neuron collects the responses from a group of afferent V4 neurons 
 and synthesizes a unified representation of shapes.
-Thus in our model,
-an IT unit combines the inputs 
-from several V4 neurons into a local image feature.
-The receptive fields of the connected V4 neurons 
-scatter around the center of the IT receptive field.
+In our model, the object recognition layer consists of IT computation units.
+Each IT unit combines the output of several V4 units into a local image feature.
+The receptive fields of these V4 units scatter around the center of the IT receptive field.
 Though there is no anatomical evidence showing that 
-the spatial distribution of the connected V4 receptive fields 
+the spatial distribution of the V4 receptive fields 
 is necessarily restricted to any fixed patterns,
-in our model, we choose afferent V4 neurons
-from a group of rectangular grid points.
-Each IT neuron has 9 afferent connections from the V4 layer.
-The 9 connected neurons are located in a $3\times 3$ grid.
-\figurename~\ref{fig:it} demonstrates an IT receptive field 
+in our model, we choose afferent V4 units from a group of rectangular grid points.
+Each IT unit has 9 afferent connections from the V4 layer.
+The 9 connected V4 units are located in a $3\times 3$ grid.
+\figurename~\ref{fig:13} demonstrates an IT receptive field 
 and the receptive fields of the connected V4 neurons.
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\linewidth]{images/fig3c.eps}
-\caption{Demonstration of IT neuron in our model. 
-IT neuron combines V4 inputs in a grid layout rotated to the strongest orientation.}
-\label{fig:it}
-\end{figure}
-
-The gird is rotated to the strongest orientation 
-of the central V4 input 
-in order to achieve the invariance of feature orientations.
-We were inspired by the SIFT \cite{Lowe1999} descriptor.
-The rotation is also demonstrated in \figurename~\ref{fig:it}.
-The grid is rotated to the orientation of the edge of the hat
-appearing in the center of the grid.
-
-The output of IT neurons are thus defined as a concatenation of the V4 input vectors.
+\begin{figure}
+\centering
+\includegraphics[width=0.5\linewidth]{images/fig-13.pdf}
+\caption{Demonstration of IT unit in our model. 
+IT unit combines V4 inputs in a grid layout rotated to the strongest orientation.}
+\label{fig:13}
+\end{figure}
+
+In order to achieve the invariance of feature orientations,
+the gird is rotated to the strongest orientation at the center of the receptive field.
+We have been inspired by the SIFT \cite{lowe1999} descriptor.
+The rotation is also demonstrated in \figurename~\ref{fig:13}.
+The grid is rotated to the orientation of the edge of the hat appearing at the center of the grid.
+The strongest central orientation can be obtained by taking a maximal value 
+over the output of the complex cells connected to the central V4 units.
+
+The output of an IT unit is thus defined as a concatenation of the V4 feature vectors.
 \begin{equation}
-\mathbf{\vec{F}}(x,y) = \mathbf{\vec{V}}(\hat{x}_1,\hat{y}_1)\|
-\mathbf{\vec{V}}(\hat{x}_2,\hat{y}_2)\|
-\mathbf{\vec{V}}(\hat{x}_3,\hat{y}_3)\|{}  \cdots\|
-\mathbf{\vec{V}}(\hat{x}_n,\hat{y}_n),
+\mathbf{F}(x,y) = \mathbf{V}(\hat{x}_1,\hat{y}_1)\|
+\mathbf{V}(\hat{x}_2,\hat{y}_2)\|
+\mathbf{V}(\hat{x}_3,\hat{y}_3)\|{}  \cdots\|
+\mathbf{V}(\hat{x}_n,\hat{y}_n),
 \end{equation}
@@
 $\hat\theta = \arg\!\max_\theta C_\theta(x,y)$.
-In this equation, $\mathbf{\vec{F}}$ represents the image feature generated by IT neurons, 
-$\hat\theta$ represents the strongest orientation of the central V4 input, 
+In this equation, $\mathbf{F}$ represents the image feature generated by IT units.
+$\hat\theta$ represents the strongest orientation at the center of the receptive field,
+which is calculated over the output of complex cells $C_\theta$.
 $(\hat{x}_i,\hat{y}_i)$ is a group of rotated grid points
-indicating the centers of receptive fields of afferent V4 neurons, 
-and $(x_i,y_i)$ is a group of grid points.
+indicating the centers of receptive fields of afferent V4 units.
+$(x_i,y_i)$ is a group of grid points.
 In our model, $(x_i,y_i)$ is defined as the $3\times 3$ grid points fitted 
@@
 
-\subsection{Object recognition}\label{sec:exp:objectrecognition}
+\subsection{Object recognition}
 
@@
 
-The bags of key-points approach \cite{Csurka2004} are employed.
-Key-points are detected with the salient region detector 
-proposed by Kadir and Brady \cite{Kadir2001}.
-Features at the key-points are extracted with our model
-and passed to a classifier.
+The bags of key-points approach \cite{csurka2004} are employed.
+Key-points are detected with the saliency sublayer of the V4 computation units in our model.
+Features at the key-points are extracted with IT units and passed to a classifier.
 We experimented with the SVM classifier
-and evaluated the classifier over the Caltech 101 dataset \cite{Fei-Fei2004}.
-The parameters of our model are summarized in \tablename~\ref{tab:param}.
-
-\begin{table}
-\caption{Model Parameters}
-\label{tab:param}
-\begin{center}
-\begin{tabular}{ll}
-\hline
-{\bf Description} &{\bf Value}\\ 
-\hline
-Radius of salient region & $R$ \\
-Grid size of IT unit ($|x_{i}-x_{i+1}|$) & $|x_{i}-x_{i+1}|=R$ \\
-Radius of complex receptive field ($\sigma_c$) & $\sigma_c=R/3$ \\
-Radius of simple receptive filed ($\sigma_s$) & $\sigma_s=\sigma_c/2$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-In addition to the parameters summarized in \tablename~\ref{tab:param},
-the sampling density of different preferred orientations $\theta_k$
-($k=1,2,3,\dots$) as in equation (\ref{equ:v4}) is also a parameter in our model.
-A density of $n$ means that $\theta_k$ is taken from the range
+and evaluated the classifier over the Caltech 101 dataset \cite{feifei2004}.
+
+The parameters that controls the size of simple, complex and V4 receptive field
+are summarized in \tablename~\ref{tab:1} in section 3.
+In addition to these parameters,
+the sampling density of different preferred orientations is also a parameter in our model.
+In the RBM encoder sublayer of V4 units,
+the visible RBM units collect their input from complex cells 
+with different preferred orientations.
+A sampling density of $n$ means that a preferred orientation is taken from the range
 $\{0^{\circ}, 180^{\circ}/n, 2\times 180^{\circ}/n, \dots, (n-1)\times 180^{\circ}/n\}$.
@@
 
-\begin{figure}[h]
+\begin{figure}
 \centering
   \subfloat[Sampling density of orientations]{
-    \includegraphics[width=0.8\linewidth]{images/fig5a.eps}
-    \label{fig:result:density}}\\
+    \includegraphics[width=0.9\linewidth]{images/fig-14a.pdf}
+    \label{fig:14a}}\\
   \subfloat[Number of features per image]{
-    \includegraphics[width=0.8\linewidth]{images/fig5b.eps}
-    \label{fig:result:nfeature}}
+    \includegraphics[width=0.9\linewidth]{images/fig-14b.pdf}
+    \label{fig:14b}}
 \caption{Performance with different parameters.}
-\label{fig:result}
-\end{figure}
-
-As shown in \figurename~\ref{fig:result:density},
-the selection of sampling density does not exhibit obvious impact
-on the performance.
-With a small density, we can obtain 
-both relatively low-dimensional feature vectors
-and satisfactory performance.
+\label{fig:14}
+\end{figure}
+
+As shown in \figurename~\ref{fig:14a},
+the selection of sampling density does not exhibit obvious impact on the performance.
+With a small density, we can lower the dimension of the input of RBM encoders
+without a significant loss in object recognition performance.
 Thus in the succeeding experiments, a sampling density of $3$ is used.
 
-To investigate the contribution of the number of features 
-on performance,
+We can also manipulate the saliency threshold in the saliency sublayer of V4 units
+to change the number of key-points detected in an image,
+which inturn changes the number of features generated by IT units.
+To investigate the contribution of the number of features on performance,
 we vary the number of features chosen from each training image.
@@
 The results reported are also averaged over 8 independent runs.
-\figurename~\ref{fig:result:nfeature} shows that 
+\figurename~\ref{fig:14b} shows that 
 with 50 features per image the result approaches the best performance.
 The number of required features can be notably reduced by introducing 
-a clustering step in the bag of words approach \cite{Csurka2004}.
-However, Serre et al. \cite{Serre2007} reported that clustering could
+a clustering step in the bag of words approach \cite{csurka2004}.
+However, Serre et al. \cite{serre2007} reported that clustering could
 be sensitive to the choice of parameters and initializations,
@@
 5,10,15,20,25 and 30.
-\figurename~\ref{fig:result:ntrain} shows the performance as
+\figurename~\ref{fig:15a} shows the performance as
 a function of the number of training images.
@@
 with 50 features extracted from each training image.
-The performance is compared with other published results
-\cite{Bosch2007,Bosch2007b,Zhang2007,Lazebnik2006,Wang2006,Grauman2005}.
-
-\begin{figure}[h]
+The performance is compared with other published results.
+
+\begin{figure}
 \centering
   \subfloat[Comparison with other methods]{
-    \includegraphics[width=0.8\linewidth]{images/fig5c.eps}
-    \label{fig:result:ntrain}}\\
+    \includegraphics[width=0.9\linewidth]{images/fig-15a.pdf}
+    \label{fig:15a}}\\
   \subfloat[Comparison with other features]{
-    \includegraphics[width=0.8\linewidth]{images/fig5d.eps}
-    \label{fig:result:compare}}
+    \includegraphics[width=0.9\linewidth]{images/fig-15b.pdf}
+    \label{fig:15b}}
 \caption{Comparison of performance.}
-\label{fig:result2}
-\end{figure}
-
-\figurename~\ref{fig:result:compare} shows
+\label{fig:15}
+\end{figure}
+
+\figurename~\ref{fig:15b} shows
 the performance of our model over the 101 categories
-compared with SIFT features \cite{Lowe1999},
-PHOW features \cite{Lazebnik2006},
-and biologically inspired HMAX model \cite{Serre2007}.
+compared with SIFT features \cite{lowe1999},
+PHOW features \cite{lazebnik2006},
+and biologically inspired HMAX model \cite{serre2007}.
 The results are obtained with 30 training images randomly selected 
@@
 The images passed to the HMAX model are resized 
-to a uniform height of 140 pixels as recommended by \cite{Serre2007}.
+to a uniform height of 140 pixels as recommended by \cite{serre2007}.
 Our model exhibits competitive performance 
 compared with other models.
-As shown in \figurename~\ref{fig:result:compare},
+As shown in \figurename~\ref{fig:15b},
 the SIFT performance points over the 101 categories

### 2015-06-29 article.tex
remove style to submit for springer

@@
-\documentclass[5p]{elsarticle}
-
-\usepackage{lineno,hyperref}
+\documentclass[10pt]{article}
+
 \usepackage{graphicx}
+\usepackage{amsmath}
 \usepackage{url}
 \usepackage{subfig}
-\usepackage{amsmath}
 \usepackage{algorithm}
@@
 \usepackage{booktabs}
-\modulolinenumbers[5]
-
-\journal{Journal of Neural Networks}
-
-\bibliographystyle{bst/elsarticle-num}
+\usepackage[sort&compress]{natbib}
 
@@
 
-\begin{frontmatter}
-
 \title{V4 Neural Network Model for Shape-based Feature Extraction and Object Discrimination}
 
-\author{Zheng Dong}
-\ead{dongzheng@fudan.edu.cn}
-
-\author{Hui Wei\corref{cor:1}}
-\cortext[cor:1]{Corresponding author}
-\ead{weihui@fudan.edu.cn}
-
-\address{Laboratory of Cognitive Modeling and Algorithms, %
-School of Computer Science, Fudan University, Shanghai, China}
+\author{Hui Wei\footnote{Corresponding author} \and Zheng Dong
+\footnote{Hui Wei and Zheng Dong are with the Laboratory of Cognitive Modeling and Algorithms, %
+School of Computer Science, Fudan University, Shanghai, China. %
+Corresponding author's email: weihui@fudan.edu.cn}}
+
+\date{}
+
+\maketitle
 
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+Visual area V4 lies in the middle of the ventral visual pathway in the primate brain. 
+It is an intermediate stage in the visual processing for object discrimination. 
+It plays an important role in the neural mechanism of visual attention and shape recognition.
+V4 neurons exhibit selectivity for salient features of contour conformation.
+In this paper, we propose a novel model of V4 neurons based on a multilayer neural network
+inspired by recent studies on V4.
+Its low-level layers consist of computational units simulating simple cells and complex cells
+in the primary visual cortex.
+These layers extract preliminary visual features including edges and orientations.
+The V4 computational units calculate the entropy of the extracted features 
+as a measure of visual saliency.
+The salient features are then selected and encoded with a layer of 
+Restricted Boltzmann Machine to generate an intermediate representation of object shapes.
+The model is evaluated in shape distinction, feature detection, feature matching 
+and object discrimination experiments.
+The results demonstrate that this model generates discriminative local representation
+of object shapes.
+It shows a successful attempt to construct computation model of the ventral pathway
+for visual object recognition.
 \end{abstract}
 
-\begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
+Keywords:
+Neural network model, visual cortex, V4, shape selectivity, feature extraction, object recognition
 
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig-1.png}} 
+\centerline{\includegraphics[width=0.8\linewidth]{images/fig-1.jpg}} 
 \caption{V4 is an intermediate stage in visual recognition.}
@@
 the model demonstrates similar selectivity compared with V4 neurons.
-and compared with other models in section~\ref{sec:exp}. 
-The conclusion is summarized in section~\ref{sec:conclusion}.
+The features extracted with the proposed model is 
+applied to object discrimination experiment in section~\ref{sec:5}
+and shows competitive performance.
+The conclusion is summarized in section~\ref{sec:6}.
 
@@
 \centering
-\subfloat[]{\includegraphics[width=0.18\linewidth]{images/fig-2a.png}\label{fig:2a}}\hfil
-\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2b.png}\label{fig:2b}}\hfil
-\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2c.png}\label{fig:2c}}
+\subfloat[]{\includegraphics[width=0.18\linewidth]{images/fig-2a.jpg}\label{fig:2a}}\hfil
+\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2b.jpg}\label{fig:2b}}\hfil
+\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2c.jpg}\label{fig:2c}}
 \caption{Shapes to examine V4 selectivity.
@@
 \begin{figure}
-\centerline{\includegraphics[width=\linewidth]{images/fig-3.pdf}} 
+\centerline{\includegraphics[width=\linewidth]{images/fig-3.eps}} 
 \caption{Multilayer neural network model of V4.}
@@
 \subfloat[Phase offset $\psi=\pi/2$]{
-  \includegraphics[width=0.4\linewidth]{images/fig-4a.pdf}
+  \includegraphics[width=0.4\linewidth]{images/fig-4a.jpg}
   \label{fig:4a}}
 \subfloat[Phase offset $\psi=0$]{
-  \includegraphics[width=0.4\linewidth]{images/fig-4b.pdf}
+  \includegraphics[width=0.4\linewidth]{images/fig-4b.jpg}
   \label{fig:4b}}
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig-5.png}} 
+\centerline{\includegraphics[width=0.8\linewidth]{images/fig-5.jpg}} 
 \caption{Entropy of image patches.
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.99\linewidth]{images/fig-6.pdf}} 
+\centerline{\includegraphics[width=0.99\linewidth]{images/fig-6.eps}} 
 \caption{RBM encoder for local shape feature.}
@@
 \centering
-\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{images/fig-7a.png}\label{fig:7a}}\hfil
-\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{images/fig-7b.png}\label{fig:7b}}
+\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{images/fig-7a.jpg}\label{fig:7a}}\hfil
+\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{images/fig-7b.jpg}\label{fig:7b}}
 \caption{Response map and weight matrices of a perceptron that distinguishes two shapes.
@@
 \centering
-\subfloat[Patterns]{\includegraphics[width=0.24\linewidth]{images/fig-8a.png}\label{fig:8a}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8b.png}\label{fig:8b}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8c.png}\label{fig:8c}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8d.png}\label{fig:8d}}
+\subfloat[Patterns]{\includegraphics[width=0.24\linewidth]{images/fig-8a.jpg}\label{fig:8a}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8b.jpg}\label{fig:8b}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8c.jpg}\label{fig:8c}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8d.jpg}\label{fig:8d}}
 \caption{RBM output neurons responding to convex curves towards certain directions.
@@
 \subfloat[Response map of a neuron selective for a pair of circular sectors (selectivity index = 3.92).]%
-{\includegraphics[width=0.9\linewidth]{images/fig-9a.png}\label{fig:9a}}\\
+{\includegraphics[width=0.9\linewidth]{images/fig-9a.jpg}\label{fig:9a}}\\
 \subfloat[Response map of a neuron selective for helix shapes (Selectivity index = 6.37).]%
-{\includegraphics[width=0.9\linewidth]{images/fig-9b.png}\label{fig:9b}}
+{\includegraphics[width=0.9\linewidth]{images/fig-9b.jpg}\label{fig:9b}}
 \caption{Response maps of RBM output neurons tuned for non-Cartesian gratings.}
@@
 \centering
-\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{images/fig-10a.png}\label{fig:10a}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10b.png}\label{fig:10b}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10c.png}\label{fig:10c}}
+\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{images/fig-10a.jpg}\label{fig:10a}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10b.jpg}\label{fig:10b}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10c.jpg}\label{fig:10c}}
 \caption{RBM output neurons responding to closed shapes.}
@@
 \centering
-\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{images/fig-11a.png}\label{fig:11a}}\hfil
-\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11c.png}\label{fig:11b}}\hfil
-\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11e.png}\label{fig:11c}}\\
-\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{images/fig-11b.png}\label{fig:11d}}\hfil
-\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11d.png}\label{fig:11e}}\hfil
-\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11f.png}\label{fig:11f}}\\
-\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11g.pdf}\label{fig:11g}}\hfil
-\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11h.pdf}\label{fig:11h}}
+\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{images/fig-11a.jpg}\label{fig:11a}}\hfil
+\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11c.jpg}\label{fig:11b}}\hfil
+\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11e.jpg}\label{fig:11c}}\\
+\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{images/fig-11b.jpg}\label{fig:11d}}\hfil
+\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11d.jpg}\label{fig:11e}}\hfil
+\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11f.jpg}\label{fig:11f}}\\
+\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11g.jpg}\label{fig:11g}}\hfil
+\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11h.jpg}\label{fig:11h}}
 \caption{Feature detection. (a) and (d) show the original images.
@@
 \section{Object discrimination}
+\label{sec:5}
 
@@
 
-\begin{figure}[h]
+\begin{figure}
 \centering
   \subfloat[Matching between the same face]{
-    \includegraphics[width=0.8\linewidth]{images/fig-12a.pdf}
+    \includegraphics[width=0.8\linewidth]{images/fig-12a.jpg}
     \label{fig:12a}}\\
   \subfloat[Matching between different faces]{
-    \includegraphics[width=0.8\linewidth]{images/fig-12b.pdf}
+    \includegraphics[width=0.8\linewidth]{images/fig-12b.jpg}
     \label{fig:12b}}
@@
 and better generalization ability compared with SIFT features.
+
+\begin{figure}
+\centering
+  \subfloat[Our match]{
+    \includegraphics[width=0.8\linewidth]{images/fig-13a.jpg}
+    \label{fig:13a}}\\
+  \subfloat[SIFT match]{
+    \includegraphics[width=0.8\linewidth]{images/fig-13b.jpg}
+    \label{fig:13b}}
+\caption{Comparison of feature matching between van images.}
+\label{fig:13}
+\end{figure}
+
+Another comparison in \figurename~\ref{fig:13} demonstrates
+more clearly the advantage of our model compared with SIFT feature.
+\figurename~\ref{fig:13a} shows that our model produced correct matches
+between the model van in the left and the two real vans in the right. 
+However, SIFT feature failed to produce correct matches (\figurename~\ref{fig:13b}). 
+SIFT feature utilizes the direction of image gradient. 
+It fails to capture the shape feature of an object when
+the object's color changes dramatically because the direction of
+image gradient changes with the color. 
+Our model focuses on the shape feature and thus achieves a stable matching despite
+the change in color.
 
@@
 \centering
-\includegraphics[width=0.5\linewidth]{images/fig-13.pdf}
+\includegraphics[width=0.5\linewidth]{images/fig-14.jpg}
 \caption{Demonstration of IT unit in our model. 
 IT unit combines V4 inputs in a grid layout rotated to the strongest orientation.}
-\label{fig:13}
+\label{fig:14}
 \end{figure}
@@
 We have been inspired by the SIFT \cite{lowe1999} descriptor.
-The rotation is also demonstrated in \figurename~\ref{fig:13}.
+The rotation is also demonstrated in \figurename~\ref{fig:14}.
 The grid is rotated to the orientation of the edge of the hat appearing at the center of the grid.
@@
   \subfloat[Sampling density of orientations]{
-    \includegraphics[width=0.9\linewidth]{images/fig-14a.pdf}
-    \label{fig:14a}}\\
+    \includegraphics[width=0.9\linewidth]{images/fig-15a.eps}
+    \label{fig:15a}}\\
   \subfloat[Number of features per image]{
-    \includegraphics[width=0.9\linewidth]{images/fig-14b.pdf}
-    \label{fig:14b}}
+    \includegraphics[width=0.9\linewidth]{images/fig-15b.eps}
+    \label{fig:15b}}
 \caption{Performance with different parameters.}
-\label{fig:14}
-\end{figure}
-
-As shown in \figurename~\ref{fig:14a},
+\label{fig:15}
+\end{figure}
+
+As shown in \figurename~\ref{fig:15a},
 the selection of sampling density does not exhibit obvious impact on the performance.
@@
 The results reported are also averaged over 8 independent runs.
-\figurename~\ref{fig:14b} shows that 
+\figurename~\ref{fig:15b} shows that 
 with 50 features per image the result approaches the best performance.
@@
 5,10,15,20,25 and 30.
-\figurename~\ref{fig:15a} shows the performance as
+\figurename~\ref{fig:16a} shows the performance as
 a function of the number of training images.
@@
   \subfloat[Comparison with other methods]{
-    \includegraphics[width=0.9\linewidth]{images/fig-15a.pdf}
-    \label{fig:15a}}\\
+    \includegraphics[width=0.9\linewidth]{images/fig-16a.eps}
+    \label{fig:16a}}\\
   \subfloat[Comparison with other features]{
-    \includegraphics[width=0.9\linewidth]{images/fig-15b.pdf}
-    \label{fig:15b}}
+    \includegraphics[width=0.9\linewidth]{images/fig-16b.eps}
+    \label{fig:16b}}
 \caption{Comparison of performance.}
-\label{fig:15}
-\end{figure}
-
-\figurename~\ref{fig:15b} shows
+\label{fig:16}
+\end{figure}
+
+\figurename~\ref{fig:16b} shows
 the performance of our model over the 101 categories
@@
 compared with other models.
-As shown in \figurename~\ref{fig:15b},
+As shown in \figurename~\ref{fig:16b},
 the SIFT performance points over the 101 categories
@@
 
-\section{Conclusion}\label{sec:conclusion}
-
-In this paper, we make an attempt 
-to construct a multi-layer neural network model 
-that explicitly maps different layers 
-to different stages in the visual pathway of object recognition. 
-We demonstrate in experiments that local image features 
-generated by this neural model show competitive performance 
-in object discrimination. 
-More over, this model provides a framework 
-to take into account higher-level stages in visual processing, 
-such as the area V4 and the IT cortex, 
-and emulates object recognition mechanisms in the cortex.
-
-There are at least two directions that could be
-followed to further improve the model.
-First, feedback connections should be
-adopted for the reason that feedback connections
-occur widely in the neural systems in the forms 
-of both intra-layer and inter-layer feedbacks.
-Second, computational units that better 
-conform to the neural response patterns should
-be well evaluated and adopted 
-to replace the current coarse implementation 
-of the V4 and IT layers.
-
-\section*{References}
-
+\section{Conclusion}\label{sec:6}
+
+In this paper, we propose a novel model for the visual area V4.
+It is based on the neural mechanism of the ventral visual pathway.
+We focus on visual attention and shape selectivity in V4.
+V4 in our model is implemented as a multilayer neural network 
+which selects salient points and encodes the shape feature in the neighborhood of the salient points.
+We demonstrate a variety of experiments in which the model was evaluated.
+The results show that the proposed model is consistent with the shape selectivity of area V4.
+It can find salient points in images and encode local shape feature into a discriminative representation.
+
+We further evaluate our model in object recognition experiment.
+We designed a image descriptor which synthesizes several V4 computation units
+into an IT unit to produce a set of invariant local image features.
+Compared with traditional image descriptors,
+the proposed model shows competitive performance in object recognition experiment. 
+It captures the essential shape feature of objects
+and thus achieves a good balance between discrimination and generalization.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{images/fig-17.jpg}
+\caption{Framework of the proposed model.}
+\label{fig:17}
+\end{figure}
+
+The framework of the proposed multilayer neural network model is shown in \figurename~\ref{fig:17}.
+Our contribution is focused on the V4 layer,
+where salient shape features are detected and located,
+and the features are encoded and further synthesized to produce discriminative representation of object shapes.
+The proposed model demonstrates a successful attempt to construct model of the ventral visual pathway
+in the brain for visual discrimination.
+We explicitly adopt the model of several layers of this pathway,
+including well studied lower-level layers, such as simple cells and complex cells.
+We focus our work on V4, which is significant but lacks mature computation models.
+
+Future work should involve improvement of the IT unit and further evaluation on learning and detection shapes.
+The implementation of the IT unit should conform to neural mechanism
+and should be capable of learning generalized representation of shape categories.
+Information about color and texture of objects is also very important,
+which should also participate in the process of object discrimination.
+These show the direction of our future work.
+
+\bibliographystyle{unsrt}
 \bibliography{ref}

### 2015-09-01 article.tex
change to doublespace
brief sec1 sec2. -- todo: add deep learning in sec2

@@
 \usepackage[sort&compress]{natbib}
+\usepackage{setspace}
 
@@
 \begin{abstract}
-Visual area V4 lies in the middle of the ventral visual pathway in the primate brain. 
-It is an intermediate stage in the visual processing for object discrimination. 
-It plays an important role in the neural mechanism of visual attention and shape recognition.
-V4 neurons exhibit selectivity for salient features of contour conformation.
-In this paper, we propose a novel model of V4 neurons based on a multilayer neural network
-inspired by recent studies on V4.
-Its low-level layers consist of computational units simulating simple cells and complex cells
-in the primary visual cortex.
+Visual area V4 plays an important role in the neural mechanism of shape recognition.
+V4 neurons exhibit selectivity for the orientation and curvature of boundary fragments.
+In this paper, we propose a novel neural network model of V4 for shape-based feature extraction and other vision tasks.
+The low-level layers of the model consist of computational units simulating simple cells and complex cells in the primary visual cortex.
 These layers extract preliminary visual features including edges and orientations.
-The V4 computational units calculate the entropy of the extracted features 
-as a measure of visual saliency.
-The salient features are then selected and encoded with a layer of 
-Restricted Boltzmann Machine to generate an intermediate representation of object shapes.
-The model is evaluated in shape distinction, feature detection, feature matching 
-and object discrimination experiments.
-The results demonstrate that this model generates discriminative local representation
-of object shapes.
-It shows a successful attempt to construct computation model of the ventral pathway
-for visual object recognition.
+The V4 computational units calculate the entropy of the extracted features as a measure of visual saliency.
+The features around salient points are then selected and encoded with a layer of Restricted Boltzmann Machine 
+to generate an intermediate representation of object shapes.
+The model is evaluated in shape distinction, feature detection, feature matching and object discrimination experiments.
+The results demonstrate that this model generates discriminative local representation of object shapes.
+It shows a successful attempt to construct a computation model of visual object recognition in the brain.
 \end{abstract}
@@
 
-Understanding the content of images has always been a difficult task in image analysis
-due to the well known semantic gap \cite{smeulders2000}.
+Understanding the content of images has always been a difficult task in image analysis due to the well known semantic gap \cite{smeulders2000}.
 A high degree of abstraction is required to transform low-level pixel data into high-level knowledge of the images.
-Computer vision researchers attempt to bridge the gap between pixel data and image content with image features.
-Various methods have been proposed to extract features from pixel representation of images.
+In an attempt to bridge this semantic gap, various methods have been proposed to extract features from the pixel representation of images.
 However, humans and primates outperform any existing computer vision system with respect to almost any measure.
-How visual cortex recognizes objects is a critical question for neuroscience.
-Visual stimuli captured by the eye are transformed into neural impulses in the retina and further processed by 
-the visual cortex for high level cognitive tasks.
-The visual cortex plays an important role in filling the gap
-between visual stimuli and the implied semantic information.
-Neural impulses travel through the visual pathway \cite{ettlinger1990}
-in the visual cortex and finally contribute to a unified percept of the object of interest.
+Though the neural mechanism of vision has not been fully revealed,
+recent research on neuroscience do provide many inspiring discoveries which would help development of computer vision algorithms.
+In this paper, we demonstrate a novel model of visual area V4 and its application to shape-based feature extraction and object discrimination.
+Experiment results show that the proposed model exhibits similar shape selectivity to that of real V4 neurons
+and it outperforms many existing methods in computer vision tasks including feature matching and image classification.
 
 Primate brains possess two distinct visual pathways \cite{ettlinger1990,lehky2007}.
-The dorsal pathway is involved with processing the object's spatial location relevant to the viewer. 
-The ventral pathway is involved with object discrimination and recognition.
+The dorsal pathway is involved with processing the object's spatial location. 
+The ventral pathway is involved with object recognition.
 In this paper, we concentrate on the latter, the ventral pathway, and visual area V4 in particular.
 V4 lies in the middle of the ventral pathway (\figurename~\ref{fig:1}).
-Lower levels of the pathway (visual areas V1 and V2) extract preliminary visual features 
-including edges and orientations \cite{hubel1962,hubel1965}.
-Higher levels of the pathway (inferior temporal cortex) exhibit selectivity to complex objects
-like faces and body parts \cite{bruce1981,bell2009}.
-As an intermediate stage, V4 plays a crucial role in transforming low-level orientation signals 
-into complex object representations.
-
-\begin{figure}
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig-1.jpg}} 
+It extracts the feature of contour conformation, which represents a middle level abstraction of visual information in the brain.
+Lower levels (V1 and V2) extract features such as edges and orientations \cite{hubel1962,hubel1965}.
+Higher levels (inferior temporal cortex) comprehend concepts of complex objects like faces and body parts \cite{bruce1981,bell2009}.
+As an intermediate stage, V4 plays a crucial role in transforming low-level orientation signals into complex object representations.
+
+\begin{figure}
+\centerline{\includegraphics[width=0.8\linewidth]{images/fig1.jpg}} 
 \caption{V4 is an intermediate stage in visual recognition.}
@@
 
-In this paper, we propose a multi-layer neural network model
-which imitates the neural mechanism of visual area V4 in the ventral visual pathway.
-Different stages of the neural pathway in the visual cortex
-are explicitly mapped to different layers of the neural network model.
+In this paper, we propose a multi-layer neural network model which imitates the neural mechanism of visual area V4 in the ventral visual pathway.
+Different stages of the neural pathway in the visual cortex are explicitly mapped to different layers of the neural network model.
 In this feed-forward network, features are extracted hierarchically from input images. 
@@
 and therefore achieve a good balance between the discrimination and generalization abilities.
-The features exhibit competitive performance when applied to object recognition tasks.
-Moreover, this multilayer neural network shows a successful attempt 
-to integrate different stages of visual processing in the brain. 
-It provides an extensible framework so that future findings in cortex mechanisms 
-can be adopted and applied to computer vision tasks.
-
-We briefly introduce some related work on traditional image features
-and the biological background of our model in section~\ref{sec:2}.
-We focus on the shape selectivity of area V4 and some previous models of V4 are also discussed in this section.
+Moreover, this multilayer neural network shows a successful attempt to integrate different stages of visual processing in the brain. 
+It provides an extensible framework so that future findings in cortex mechanisms can be adopted and applied to computer vision tasks.
+
+We briefly introduce some related work on image features and V4 models in section~\ref{sec:2}.
 In section~\ref{sec:3}, we give the detailed description of our proposed model.
-It is capable of locating salient points 
-and generating discriminative local shape features around each salient point.
-In section~\ref{sec:4}, empirical evaluation shows that
-the model demonstrates similar selectivity compared with V4 neurons.
-The features extracted with the proposed model is 
-applied to object discrimination experiment in section~\ref{sec:5}
-and shows competitive performance.
-The conclusion is summarized in section~\ref{sec:6}.
+It is capable of locating salient points and generating discriminative local shape features around salient points.
+In section~\ref{sec:4}, we show several experiments in which we evaluate the proposed model.
+The model is compared with real V4 neurons in a shape distinction experiment.
+The features extracted by the model are applied to various computer vision tasks.
+The conclusion is summarized in section~\ref{sec:5}.
 
@@
 
+In this section, we briefly explain the difference between our model and traditional image features.
+We also review several existing models of V4 and some other related models.
+
 \subsection{Traditional image features}
 
-Image features encode the image structure in a spatial neighborhood around each of a set of interest points.
-Extracting features is an important stage in the process of understanding the content of an image.
-These features can be low-level features, such as edges \cite{canny1986} 
-and corners \cite{harris1988,smith1997},
-or high-level features, which are usually more distinctive and descriptive.
-
-Traditional image features mostly fall into three categories \cite{mikolajczyk2005}: 
+The model proposed in this paper is a biologically motivated model for image feature extraction.
+As reviewed in \cite{mikolajczyk2005}, traditional image features fall into three categories: 
 distribution-based descriptors, spatial-frequency descriptors and differential-based descriptors.
-Distribution-based descriptors use histograms 
-to represent characteristics of local image regions.
-SIFT descriptor \cite{lowe1999} (including its variations) 
-and shape context \cite{belongie2002} are typical examples of such descriptors.
-SIFT descriptor uses histograms 
-of oriented image gradients over rectangular grids.
-Shape context implements similar idea 
-by computing histograms of edge pixels in a log-polar grid.
-Spatial-frequency descriptors and differential-based descriptors 
-use filters and derivatives to capture the invariant features. 
-Examples of these categories are 
-complex filters \cite{schaffalitzky2002} 
-and Baumberg's affine invariant descriptor \cite{baumberg2000}.
-Complex filters are a family of filters 
-derived by adding complex multipliers to a Gaussian filter.
-Baumberg's descriptor is generated 
-by applying an elliptical Gaussian filter 
-to the second order moment 
-of the scale-space derivative \cite{lindeberg1993} of images.
-
-These traditional image features utilize flat models which compute features directly from the image
-via statistical or algebraic techniques. 
-The flat approach does not conform to the hierarchical mechanism \cite{ungerleider2011} 
-employed by neural systems such as the primate brain.
-
-\subsection{Neural mechanism of vision}
-
-At lower levels of the ventral visual pathway (V1 and V2), neurons respond to features of local orientation. 
-These neurons fall into two categories, 
-simple cells and complex cells \cite{hubel1962,hubel1965,martinez2003}.
-The receptive fields of simple cells 
-can be understood as linear filters modeled as Gabor functions \cite{gabor1946}. 
-They respond primarily to oriented edges and gratings.
-Complex cells differ from simple cells in that a stimulus is
-effective wherever it is placed in the receptive field, provided
-that the orientation is appropriate \cite{hubel1962,hubel1965}.
-
-Neurons in V4 and IT cortex encode high-level information.
-The exact nature of the encoded information remains to be determined.
-The response pattern of V4 neurons shows 
-a strong bias towards convex contour features \cite{pasupathy1999,pasupathy2001,pasupathy2002}. 
-V4 neurons are selective for orientation and curvature of boundary conformation.
-Research also suggests that simple shapes are represented 
-by patterns of activity across populations of V4 neurons \cite{pasupathy2002}.
-
-IT cortex is significant in visual recognition and
-discrimination \cite{desimone1980,bell2009,ungerleider2011,bruce1981}.
-It depends exclusively on inputs received from
-connections originating in V1 \cite{desimone1980}.
-Recently, object representations in the temporal cortex of monkeys and humans are revealed 
-by functional magnetic resonance imaging \cite{bell2009}.
-Discrete regions selective for animate and inanimate stimuli were found in both species.
-These regions could be further subdivided into regions selective for individual categories.
-
-\subsection{Shape Selectivity of V4 Neurons}
-
-Neurobiological studies on V4 have not produced a unified model of its function or circuitry.
-V4 neurons are known to be selective for color, shape, depth and even motion \cite{roe2012}.
-In this paper, we focus on the V4 selectivity for shapes. 
-Early experiments examined the selectivity of cells in V4 
-with classical stimuli including bars and sinusoidal gratings~(\figurename~\ref{fig:2a}) \cite{desimone1987}.
-Similar to earlier processing stages, some V4 neurons are tuned for orientation 
-and spatial frequency of edges and linear sinusoidal gratings.
-However, the majority of V4 neurons are sensitive to more complex shape properties. 
-Later experiments demonstrated that V4 neurons display a clear bias in their responses 
-in favor of non-Cartesian gratings~(\figurename~\ref{fig:2b}) and they show a significant degree of invariance 
-in their selectivity across changes in stimulus position \cite{gallant1996}.
-More recent experiments showed that V4 neurons can be strongly selective for curvature
-of contours and angular position of acute curvatures \cite{pasupathy1999,pasupathy2001}. 
-\figurename~\ref{fig:2c} shows a response map of a V4 neuron
-(reproduced from \cite{pasupathy2001}).
-The white shapes in \figurename~\ref{fig:2c} is presented in the receptive field of this V4 neuron
-with equally dark background. 
-The gray scale in the response map indicates the strength of the neuronal response.
-Darker background indicates that the neuronal response is stronger.
-This neuron is selectively tuned for acute convex border at the bottom right.
-The curvature and the angular position are both necessary conditions of the activation of this neuron.
-Neither rounded protrusions nor sharp curvatures towards directions other than the bottom left
-activate the neuron.
-
-\begin{figure}
-\centering
-\subfloat[]{\includegraphics[width=0.18\linewidth]{images/fig-2a.jpg}\label{fig:2a}}\hfil
-\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2b.jpg}\label{fig:2b}}\hfil
-\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig-2c.jpg}\label{fig:2c}}
+Distribution-based descriptors use histograms to represent characteristics of local image regions.
+SIFT descriptor \cite{lowe1999} and shape context \cite{belongie2002} are typical examples of such descriptors.
+Spatial-frequency descriptors and differential-based descriptors use filters and derivatives to capture the invariant features. 
+Examples of these categories are complex filters \cite{schaffalitzky2002} and Baumberg's affine invariant descriptor \cite{baumberg2000}.
+These traditional image features utilize flat models which compute features directly from the image via statistical or algebraic techniques. 
+They have little in common with the neural systems.
+On the contrary, our proposed model simulates the biological vision in the level of neural mechanism.
+It is a multilayer model which employs the hierarchical structure of neural systems and extracts information from simple features to gradually complex features.
+
+\subsection{Previous V4 models}
+
+V4 neurons have been examine with different varieties of visual stimuli, including light bars, sinusoidal gratings~(\figurename~\ref{fig:2a}),
+non-Cartesian gratings~(\figurename~\ref{fig:2b}), closed shapes~(\figurename~\ref{fig:2c}) \cite{desimone1987,gallant1996,pasupathy1999,pasupathy2001}.
+They generally prefer complex features to simple features such as bars and sinusoidal gratings \cite{gallant1996}.
+Most V4 neurons are tuned for the curvature and orientation of boundary fragments \cite{pasupathy2001}.
+Therefore, unlike neurons in V1, V4 neurons can not be modeled with linear filters.
+
+\begin{figure}
+\centering
+\subfloat[]{\includegraphics[width=0.18\linewidth]{images/fig2a.jpg}\label{fig:2a}}\hfil
+\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig2b.jpg}\label{fig:2b}}\hfil
+\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig2c.jpg}\label{fig:2c}}
 \caption{Shapes to examine V4 selectivity.
-(a) Classical gratings.
-(b) Non-Cartesian gratings. V4 neurons prefer non-Cartesian gratings rather than classical gratings.
-(c) Response map of a V4 neuron which responds to a sharp convex curvature at the bottom left.
-Darker background indicates a stronger response.}
+(a) Sinusoidal gratings. (b) Non-Cartesian gratings. (c) Closed shapes.}
 \label{fig:2}
 \end{figure}
-
-\subsection{Previous models of V4}
 
@@
 We briefly introduce the SRF model \cite{david2006} and the HMAX model \cite{riesenhuber1999,cadieu2007}.
-
-The spectral receptive field (SRF) \cite{david2006}
-describes properties of V4 receptive field in terms of the orientation and spatial frequency spectrum.
-The model is based on the fact that V4 neurons have large orientation and spatial frequency bandwidth.
-They respond selectively to stimuli such as contour conformations and non-Cartesian gratings,
-which generally consist of multiple orientations and spatial frequencies.
-The spectral model is also invariant to small changes in stimulus position 
-and thus explains the invariance property of V4 response patterns.
-The model is powerful in describing the shape selectivity of V4 neurons.
-However, it does not explain the emergence of the selectivity.
-It does not either provide the neural computing process to achieve such spectral receptive field.
-
-The HMAX model was first proposed in \cite{riesenhuber1999}
-as a generic model for object recognition in the visual cortex.
-It models the visual cortex into a hierarchical architecture 
-consisting of cascaded linear filters and non-linear maximum pooling operations.
-It was then adopted as a model for V4 shape selectivity and invariance \cite{cadieu2007}.
+The spectral receptive field (SRF) \cite{david2006} describes properties of V4 receptive field in terms of the orientation and spatial frequency spectrum.
+This model is powerful in describing the shape selectivity of V4 neurons.
+However, it does not provide any computational or neural mechanism to achieve V4 selectivity via spectral receptive field.
+The HMAX model was first proposed in \cite{riesenhuber1999} as a generic model for object recognition in the visual cortex.
+It models the visual cortex as a hierarchical architecture consisting of cascaded linear filters and non-linear maximum pooling operations.
+It has been also adopted as a model for V4 shape selectivity and invariance \cite{cadieu2007}.
 The training of the model is an NP-complete problem. 
-The authors used a greedy algorithm to obtain approximated solutions
-but they did not provide any biological evidence for the algorithm.
-
-The proposed model of V4 in this paper overcomes the limitation of the previous models.
-In the next section, we demonstrate that its architecture and function is analogous 
-to those of the visual cortex.
-We also provide efficient training method for our model.
+The authors used a greedy algorithm to obtain approximated solutions but they did not provide any biological evidence for the algorithm.
 

### 2015-09-02 article.tex
brief sec 1 and sec2, deep learning added

@@
 
-Primate brains possess two distinct visual pathways \cite{ettlinger1990,lehky2007}.
+Primate brains possess two distinct visual pathways \cite{ettlinger1990}.
 The dorsal pathway is involved with processing the object's spatial location. 
@@
 It extracts the feature of contour conformation, which represents a middle level abstraction of visual information in the brain.
-Lower levels (V1 and V2) extract features such as edges and orientations \cite{hubel1962,hubel1965}.
-Higher levels (inferior temporal cortex) comprehend concepts of complex objects like faces and body parts \cite{bruce1981,bell2009}.
+Lower levels (V1 and V2) extract features such as edges and orientations \cite{hubel1962}.
+Higher levels (inferior temporal cortex) comprehend concepts of complex objects like faces and body parts \cite{bell2009}.
 As an intermediate stage, V4 plays a crucial role in transforming low-level orientation signals into complex object representations.
@@
 High-level local image features emerge in the output layer of the network. 
-The features encode local structures of object shapes in a way which closely resembles the mechanism of the cortex 
-and therefore achieve a good balance between the discrimination and generalization abilities.
+The features encode local structures of object shapes in a way which closely resembles the mechanism of the cortex.
+With a limited number of training images, the model can learn to distinguish between different categories of objects.
 Moreover, this multilayer neural network shows a successful attempt to integrate different stages of visual processing in the brain. 
@@
 The features extracted by the model are applied to various computer vision tasks.
-The conclusion is summarized in section~\ref{sec:5}.
+The discussion and conclusion is summarized in section~\ref{sec:5}.
 
@@
 
-In this section, we briefly explain the difference between our model and traditional image features.
-We also review several existing models of V4 and some other related models.
-
-\subsection{Traditional image features}
-
-The model proposed in this paper is a biologically motivated model for image feature extraction.
-As reviewed in \cite{mikolajczyk2005}, traditional image features fall into three categories: 
-distribution-based descriptors, spatial-frequency descriptors and differential-based descriptors.
-Distribution-based descriptors use histograms to represent characteristics of local image regions.
-SIFT descriptor \cite{lowe1999} and shape context \cite{belongie2002} are typical examples of such descriptors.
-Spatial-frequency descriptors and differential-based descriptors use filters and derivatives to capture the invariant features. 
-Examples of these categories are complex filters \cite{schaffalitzky2002} and Baumberg's affine invariant descriptor \cite{baumberg2000}.
-These traditional image features utilize flat models which compute features directly from the image via statistical or algebraic techniques. 
-They have little in common with the neural systems.
-On the contrary, our proposed model simulates the biological vision in the level of neural mechanism.
-It is a multilayer model which employs the hierarchical structure of neural systems and extracts information from simple features to gradually complex features.
-
-\subsection{Previous V4 models}
+In this section, we briefly review some existing models of V4 and other related models.
 
 V4 neurons have been examine with different varieties of visual stimuli, including light bars, sinusoidal gratings~(\figurename~\ref{fig:2a}),
-non-Cartesian gratings~(\figurename~\ref{fig:2b}), closed shapes~(\figurename~\ref{fig:2c}) \cite{desimone1987,gallant1996,pasupathy1999,pasupathy2001}.
+non-Cartesian gratings~(\figurename~\ref{fig:2b}), and closed shapes~(\figurename~\ref{fig:2c}) \cite{gallant1996,pasupathy2001}.
 They generally prefer complex features to simple features such as bars and sinusoidal gratings \cite{gallant1996}.
@@
 Several models have been proposed for V4 neurons.
-We briefly introduce the SRF model \cite{david2006} and the HMAX model \cite{riesenhuber1999,cadieu2007}.
+We briefly introduce the SRF model \cite{david2006} and the HMAX model \cite{cadieu2007}.
 The spectral receptive field (SRF) \cite{david2006} describes properties of V4 receptive field in terms of the orientation and spatial frequency spectrum.
@@
 It has been also adopted as a model for V4 shape selectivity and invariance \cite{cadieu2007}.
-The training of the model is an NP-complete problem. 
+Training this model is an NP-complete problem. 
 The authors used a greedy algorithm to obtain approximated solutions but they did not provide any biological evidence for the algorithm.
+
+The deep learning model also draws inspiration from the layered structure of the cortex \cite{bengio2009}.
+It learns representations of data through stacked layers of Restricted Boltzmann Machines (RBM).
+When it is trained with natural images, 
+the kernels of the data-connected layer resemble the pattern of the receptive field of neurons in the primary visual cortex.
+By training many stacked layers of RBM, highly abstracted features can be extracted automatically from input images.
+Therefore, with additional fine tuning, the deep model exhibits high performance in image classification \cite{krizhevsky2012}.
+Our proposed model employs a similar multilayer structure but combines automatic feature extraction with artificial features.
+This combined strategy allows our model to work on fewer training images.
 
@@
 These areas have been well studied since 1960s 
-by Hubel, Wiesel \cite{hubel1962,hubel1965} and succeeding researchers.
+by Hubel, Wiesel \cite{hubel1962} and succeeding researchers.
 
@@
 
+\subsection{Traditional image features}
+
+The model proposed in this paper is a biologically motivated model for image feature extraction.
+As reviewed in \cite{mikolajczyk2005}, traditional image features fall into three categories: 
+distribution-based descriptors, spatial-frequency descriptors and differential-based descriptors.
+Distribution-based descriptors use histograms to represent characteristics of local image regions.
+SIFT descriptor \cite{lowe1999} and shape context \cite{belongie2002} are typical examples of such descriptors.
+Spatial-frequency descriptors and differential-based descriptors use filters and derivatives to capture the invariant features. 
+Examples of these categories are complex filters \cite{schaffalitzky2002} and Baumberg's affine invariant descriptor \cite{baumberg2000}.
+These traditional image features utilize flat models which compute features directly from the image via statistical or algebraic techniques. 
+They have little in common with the neural systems.
+On the contrary, our proposed model simulates the biological vision in the level of neural mechanism.
+It is a multilayer model which employs the hierarchical structure of neural systems and extracts information from simple features to gradually complex features.
+
+This process requires a large number of training samples (\cite{krizhevsky2012} uses 1.2 million training images)
+and a neural network of very large scale (In \cite{krizhevsky2012}, the network has 8 layers and more than 400,000 neurons).
+Deep learning can be viewed as automatic feature extraction.
+In our proposed model, we combine automatic feature extraction with artificial feature extraction.
+With prior knowledge of the cortical neurons, we manually design the layers corresponding to the primary visual cortex
+and V4 features are automatically extracted based on the output of the artificially designed network.
+This strategy saves us from retraining the first layers every time
+and it allows the model to work on limited number of training images.
+
 In this paper, we propose a novel model for the visual area V4.

### 2015-09-04 article.tex
about to finish 3.2

@@
 Our model employs a similar hierarchical structure.
-Different cortex areas along the visual pathway are explicitly adopted in our model. 
+Different cortical areas along the neural pathway are explicitly mapped to different layers in our model. 
 \figurename~\ref{fig:3} shows the architecture of our model.
 
-\subsection{Innovation in our model}
-
-The application of neural science in computer vision has been very limited.
-The most widely used DoG filters and Gabor filters 
-are models of ganglion cells in the retina and simple cells in the primary visual cortex, 
-which correspond to early stages of visual processing in the brain.
-
-Biologically inspired models \cite{serre2007, hong2011},
-proposed recently have exhibited competitive performance in traditional computer vision tasks.
-However, these models are still limited to the utilization 
-of the structure of the primary visual cortex. 
-Most of these models originate from the HMAX \cite{riesenhuber1999} model 
-which duplicates the structure of simple cells and complex cells
-in the primary visual cortex for high-level processing.
-
-In our model, we attempt to map each stage in the ventral pathway of visual processing
-to a single layer of computational units.
-These units are carefully designed so that they conform to the neural mechanism.
-We focus on the neural mechanism of V4 with particular concern
-because the shape features extracted in V4 are significant in the process of object discrimination.
-
 \begin{figure}
-\centerline{\includegraphics[width=\linewidth]{images/fig-3.eps}} 
+\centerline{\includegraphics[width=\linewidth]{images/fig3.eps}} 
 \caption{Multilayer neural network model of V4.}
@@
 
-\subsection{Layer of simple cells}
+\subsection{Layer of simple cells and complex cells}
 
-According to the hierarchy of the ventral visual pathway,
-area V4 receives input from the lower levels including area V1 and V2.
-These areas have been well studied since 1960s 
-by Hubel, Wiesel \cite{hubel1962} and succeeding researchers.
-
+According to the hierarchy of the ventral visual pathway, area V4 receives input from the lower levels including area V1 and V2.
 Neurons in V1 and V2 respond to local orientations.
 They fall into two categories, simple cells and complex cells.
-Simple cells respond primarily to oriented edges and gratings.
+A simple cell is activated when an edge stimulus is presented in its receptive field with appropriate orientation and position.
 Complex cells have larger receptive fields.
-A stimulus is effective wherever it is placed in the complex receptive field, 
-provided that the orientation is appropriate \cite{hubel1962}.
+A stimulus is effective wherever it is placed in the complex receptive field, provided that the orientation is appropriate \cite{hubel1962}.
 
 In our model, the layer of simple cells covers the whole dimension of input images.
-For each location, several simple cells with different preferred orientations 
-detect edges of different orientations.
-Simple cells are modeled as Gabor functions \cite{gabor1946},
+For each location, several simple cells with different preferred orientations detect edges of different orientations.
+Simple cells are modeled as Gabor functions,
 \begin{equation}\label{equ:gabor}
@@
 \subfloat[Phase offset $\psi=\pi/2$]{
-  \includegraphics[width=0.4\linewidth]{images/fig-4a.jpg}
+  \includegraphics[width=0.4\linewidth]{images/fig4a.jpg}
   \label{fig:4a}}
 \subfloat[Phase offset $\psi=0$]{
-  \includegraphics[width=0.4\linewidth]{images/fig-4b.jpg}
+  \includegraphics[width=0.4\linewidth]{images/fig4b.jpg}
   \label{fig:4b}}
@@
 \figurename~\ref{fig:4}~shows two Gabor functions with different phase offsets.
-The sinusoidal factors of the two Gabor functions 
-are odd function and even function respectively.
-The odd Gabor function resembles a gradient operator
-while the even Gabor function resembles a Laplacian operator.
-In a computational point of view, 
-the Laplacian operator calculates a higher order derivative 
-than the gradient operator.
+The sinusoidal factors of the two Gabor functions are odd function and even function respectively.
+The odd Gabor function resembles a gradient operator while the even Gabor function resembles a Laplacian operator.
+In a computational point of view, the Laplacian operator calculates a higher order derivative than the gradient operator.
 The gradient operator suppresses the smooth area in images.
-Moreover, the Laplacian operator is able 
-to suppress area that varies smoothly, 
-which is also neglected by human eyes generally.
-Therefore, in our model, the even Gabor function 
-with the phase offset $\psi=0$ is employed.
+Moreover, the Laplacian operator is able to suppress area that varies smoothly, which is also neglected by human eyes generally.
+Therefore, in our model, the even Gabor function with the phase offset $\psi=0$ is employed.
 
 The layer of simple cells operates on raw image input. 
-The output of simple cells with preferred orientation $\theta$
-and scale $\sigma_s$ is the following convolution passed through a transfer function $\phi$,
+The output of simple cells with preferred orientation $\theta$ and scale $\sigma_s$ is the following convolution passed through a transfer function $\phi$,
 \begin{equation}\label{equ:gabor}
@@
 
-\subsection{Layer of complex cells}
-
 Complex cells are modeled as Gaussian filters which take inputs from simple cells,
@@
 where $\sigma_c$ represents the radius of the complex receptive fields.
-The filters are discretized into matrices and normalized before applied to digital images.
 The radius of complex receptive field is approximately twice as long as that of a simple cell.
 We take this size because qualitatively the complex receptive fields are usually two to five times wider 
-than the bars of gratings that stimulate them most effectively \cite{movshon1978}.
+than the bars of gratings that stimulate them most effectively \cite{hubel1962}.
 In this paper, we have $\sigma_c=2\sigma_s$.
@@
 \begin{equation}
-C_{\theta}=|S_{\theta}|\otimes f.
+C_{\theta}=S_{\theta}\otimes f.
 \label{equ:complex}
 \end{equation}
-We use absolute values of simple cell outputs 
-because we do not distinguish between dark edges and bright edges.
-Generally speaking, simple cells modeled as in equation (\ref{equ:gabor}) 
-with phase offset $\psi=0$ response best to stimuli of bright stripes in dark background, i.e., bright edges.
-Such cells are thus called ON-center \cite{hubel1962} cells in the neural system.
-Dark edges result in a negative response (strong inhibition) of ON-center cells.
-In the neural system, OFF-center cells with inversed receptive fields response to dark edges.
-In our model, ON-center and OFF-center cells are not distinguished.
-In other words, the layer of complex cells in our model is sensitive to bright edges, dark edges 
-and boundaries between bright and dark areas.
+
+In the HMAX model \cite{riesenhuber1999} and the deep convolutional neural network \cite{krizhevsky2012}, pooling operation is an important step.
+Pooling features over a local neighborhood can create invariance to small transformations of the input \cite{boureau2010}.
+In our model, complex cells carry out similar function.
+The operation can be viewed as a Gaussian-weighted average pooling.
+We choose weighted average pooling because average pooling can be efficiently calculated with artificial neurons
+and weighted average gives the complex cell's receptive field a smooth and circular boundary.
 
@@
 
-V4 is an area of attentional modulation \cite{roe2012}.
 Visual attention involves selecting an interested region or selecting specific object features.
-Visual attention in V4 is influenced by both top-down feedback from higher levels in the visual pathway
-and bottom-up input from lower levels.
-We focus on the bottom-up influence.
-In a bottom-up process, V4 evaluates the saliency of the input from lower levels
-and focuses its attention automatically on the salient features.
-
-We use entropy to measure the saliency of images \cite{kadir2001}.
-In \cite{kadir2001}, it is assumed that salient regions have high complexity (and correspondingly high entropy).The entropy of features is used as a scale invariant measure.
-The salient region is selected at entropy peaks over scales.
-To avoid erroneously selecting noise or texture as salient regions,
-a measure of self-similarity is employed.
-Self-similar regions are filtered out.
-
-In this paper, we use entropy in a different approach.
-V4 neurons encode fragments of object contour \cite{pasupathy2001,pasupathy2002}.
-They are selective for simple structures such as convex or concave curvature.
-Therefore, in our assumption, 
-well ordered structures with low complexity are preferred in cognitive activities.
-We filter out regions with high complexity (or entropy).
-
+Area V4 shows selective attention based on spatial location \cite{desimone1995}.
+The underlying neural mechanism is unknown.
+V4 neurons exhibit preference for geometrically ordered stimuli \cite{sasaki2005}.
+Therefore we use entropy (a measure of disorder) to measure the saliency of images.
 The entropy is calculated according to the output of complex cells.
@@
 a complex cell with preferred orientation $\theta$ has output value $C_{\theta}(x,y)$,
-We suppose that the probability of a complex cell being activated
-is proportional to the output value.
+We suppose that the probability of a complex cell being activated is proportional to the output value.
 The probability is thus defined as:
@@
 \end{equation}
-The entropy of the complex cell activity in this neighborhood
-is 
+The entropy of the complex cell activity in this neighborhood is 
 \begin{equation}
@@
 \figurename~\ref{fig:5} shows four image patches.
-For each image patch, the output values of complex cells with different preferred orientations
-(from $0^\circ$ to $180^\circ$) are plotted in a bar chart.
+For each image patch, the output values of complex cells with different preferred orientations (from $0^\circ$ to $180^\circ$) are plotted in a bar chart.
 The charts show the distribution of complex cell activities.
-The entropy calculated accordingly indicates that 
-patches composed of simple structures have non-uniform distributions and thus low entropy values.
+The entropy calculated accordingly indicates that patches composed of simple structures have non-uniform distributions and thus low entropy values.
 Therefore, we filter out regions with high entropy.
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig-5.jpg}} 
+\centerline{\includegraphics[width=0.8\linewidth]{images/fig5.jpg}} 
 \caption{Entropy of image patches.
@@
 
-In addition to entropy, 
-local competition also plays a role in attentional selection \cite{desimone1995}.
+In addition to entropy, local competition also plays a role in attentional selection \cite{desimone1995}.
 With limited neural resources, only strong and competitive neuronal signals get transmitted and processed.

### 2015-09-05 article.tex
finish sec 3

@@
-\documentclass[10pt]{article}
+\documentclass[9pt,twocolumn]{article}
 
@@
 
-Primate brains possess two distinct visual pathways \cite{ettlinger1990}.
-The dorsal pathway is involved with processing the object's spatial location. 
-The ventral pathway is involved with object recognition.
-In this paper, we concentrate on the latter, the ventral pathway, and visual area V4 in particular.
-V4 lies in the middle of the ventral pathway (\figurename~\ref{fig:1}).
+V4 lies in the middle of the ventral visual pathway \cite{ettlinger1990}, 
+which is the neural stream involved with object recognition in the brain (\figurename~\ref{fig:1}).
 It extracts the feature of contour conformation, which represents a middle level abstraction of visual information in the brain.
@@
 
-In this paper, we propose a multi-layer neural network model which imitates the neural mechanism of visual area V4 in the ventral visual pathway.
+In this paper, we propose a multilayer neural network model which imitates the neural mechanism of visual area V4 in the ventral visual pathway.
 Different stages of the neural pathway in the visual cortex are explicitly mapped to different layers of the neural network model.
@@
 
-\section{Multi-layer neural network model}\label{sec:3}
+\section{Multilayer neural network model}\label{sec:3}
 
@@
 
-\subsection{Layer of simple cells and complex cells}
+\subsection{Layers of simple cells and complex cells}
 
@@
 The algorithm is listed in Algorithm \ref{alg:1}.
+In our experiment, the saliency filters take a threshold $t_E=2.8$ for the entropy
+and a threshold $t_A=0.4$ for complex cell output value (or complex cell activity).
+These values are roughly the median values of natural images.
 
@@
 
-With saliency filters described in the previous subsection,
-we are able to focus on a limited number of salient points.
+With saliency filters described in the previous subsection, we are able to focus on a limited number of salient points.
 The V4 computation units in our model encode the shape in the neighborhood of each salient point.
 The encoding is achieved with Restricted Boltzmann Machine (RBM).
-
-RBM can learn a probability distribution over its set of inputs.
-It has been found efficient in training deep neural network \cite{hinton2006}.
-The encoder layer in our model is part of a deep network.
-Therefore, we use the RBM as a training model for this encoder layer.
-The RBM is trained to encode shape features in the neighborhood of a salient point.
-\figurename~\ref{fig:6} shows such an RBM encoder.
-
-Let $(x,y)$ denote a salient point.
-For each preferred orientation, 
-we select the complex cell of which the receptive field is centered at point $(x,y)$,
-and its eight neighbors.
+RBM is widely used in unsupervised training of deep neural network \cite{bengio2009}.
+\figurename~\ref{fig:6} shows such an RBM encoder in our model.
+It encodes local shape features around a salient point, $(x,y)$.
+For each preferred orientation, we select the complex cell of which the receptive field is centered at point $(x,y)$, and its eight neighbors.
 \figurename~\ref{fig:6} demonstrates an example of three preferred orientations.
-In this case, $9\times3$ complex cells are selected.
-The number of selected neurons depends on the number of orientations we choose.
-The selected complex cells are then used as the input layer (or visible layer) of the RBM.
-The representation of the shape in the neighborhood of $(x,y)$
-is formed in the output layer (or hidden layer) of the RBM.
+In this case, $9\times3$ complex cells are selected to form the input layer (or visible layer) of the RBM.
+The output vector of the hidden layer forms a representation of local shape features.
 
 \begin{figure}
-\centerline{\includegraphics[width=0.99\linewidth]{images/fig-6.eps}} 
+\centerline{\includegraphics[width=0.99\linewidth]{images/fig6.eps}} 
 \caption{RBM encoder for local shape feature.}
@@
 
-RBM can be trained efficiently with a contrastive divergence learning algorithm \cite{hinton2002}.
-Let $w_{ij}$ be the weight of the connection from the $i$-th visible unit to the $j$-th hidden unit.
+We followed \cite{hinton2010} for the training algorithm of RBM and the choice of parameters.
+Weight-decay is used to reduce overfitting by adding an L2 penalty term, $\frac{1}{2}\lambda w^2$, to the energy function.
+Let $w_{ij}$ be the connection weight between the $i$-th visible unit and the $j$-th hidden unit.
 In each learning iteration, the change in the weight is given by
-\begin{equation}
-\Delta w_{ij}=\epsilon\left(\langle v_i h_j\rangle_\text{data}-\langle v_i h_j\rangle_\text{recon}\right),
-\end{equation}
-where $\epsilon$ is a learning rate, 
-$\langle v_i h_j\rangle_\text{data}$ is the product of two units when the visible layer is given real data,
-and $\langle v_i h_j\rangle_\text{recon}$ is the product of two units 
-when the visible layer is given reconstructed data.
-
-We use weight-decay \cite{hinton2010}
-to reduce overfitting by adding an L2 penalty term, $\frac{1}{2}\lambda w^2$.
-The weight change is then given by
 \begin{equation}
@@
 \end{equation}
-where $\lambda$ is a weight-cost coefficient.
-We followed \cite{hinton2010} for the choice of the coefficient $\lambda$ and the learning rate $\epsilon$
-in our experiments.
-
-The output vector of the hidden layer forms a representation of local shape features.
-It can be used directly as input for classification tasks.
-We can also use the RBM connection weight matrix to initialize a multilayer neural network
-for supervised back-propagation training.
-
-\subsection{Model Parameters}
-
-The size of the receptive field increases along the hierarchy of the visual system.
-Lower levels have relatively small receptive fields while higher levels have larger receptive fields.
-The size of receptive field in our model follows the same scheme.
-The relationship is shown in Table \ref{tab:1}.
-V4 neurons receive afferent connections from $3\times3$ complex cells
-that have partially overlapped receptive fields.
-Therefore the radius of the V4 receptive field is not 3 times as long as that of the complex cell.
-
-\begin{table}[h]
-\caption{Size of receptive field}
-\centering
-\begin{tabular}{rr}
-\toprule
-Category & Radius of receptive field \\
-\midrule
-Simple cells & $\sigma_s$ \\
-Complex cells & $\sigma_c = 2\sigma_s$ \\
-V4 RBM encoders & $2.5\sigma_c$\\
-\bottomrule
-\end{tabular}
-\label{tab:1}
-\end{table}
-
-We used two schemes for the number of different preferred orientations of simple cells and complex cells.
-For black and white images of shapes and handwritten digits in our experiment,
-we used 4 orientations (from $0^\circ$ to $135^\circ$ in steps of $45^\circ$).
-For gray scale images, we used 18 orientations (from $0^\circ$ to $170^\circ$ in steps of $10^\circ$)
-in order to preserve more information.
-
-The output of complex cells was normalized to the $[0,1]$ interval
-to serve as the input of RMB encoders.
-The saliency filters took a threshold $t_E=2.8$ for the entropy
-and a threshold $t_A=0.4$ for complex cell output value (or complex cell activity).
-These values were roughly the median values of natural images.
+where $\epsilon$ is the learning rate, $\lambda$ is the weight-cost coefficient,
+$\langle v_i h_j\rangle_\text{data}$ is the product of two units when the visible layer is given real data,
+and $\langle v_i h_j\rangle_\text{recon}$ is the product of two units when the visible layer is given reconstructed data.
 

### 2015-09-05 article.tex
finish experiment part.

@@
-\documentclass[9pt,twocolumn]{article}
+\documentclass[twocolumn]{article}
 
@@
 Experiment results show that the proposed model exhibits similar shape selectivity to that of real V4 neurons
-and it outperforms many existing methods in computer vision tasks including feature matching and image classification.
+and it outperforms many existing methods in computer vision tasks including feature matching and object recognition.
 
@@
 
-According to the hierarchy of the ventral visual pathway, area V4 receives input from the lower levels including area V1 and V2.
+According to the hierarchy of the visual cortex, area V4 receives input from the lower levels including area V1 and V2.
 Neurons in V1 and V2 respond to local orientations.
@@
 
-In our model, the layer of simple cells covers the whole dimension of input images.
-For each location, several simple cells with different preferred orientations detect edges of different orientations.
-Simple cells are modeled as Gabor functions,
+In our model, simple cells are modeled as Gabor functions,
 \begin{equation}\label{equ:gabor}
@@
 =\exp \left(-\frac{x'^2+y'^2}{2\sigma_s^2}\right)
-\cos \left(2\pi\frac{x'}{\lambda}+\psi\right),
+\cos\frac{2\pi x'}{\lambda},
 \end{equation}
@@
 In the equation, $\lambda$ is the wavelength of the sinusoidal factor, 
-$\psi$ is the phase offset, 
-$\theta$ represents the preferred orientation and 
-$\sigma_s$ approximates the radius of the receptive fields.
+$\theta$ represents the preferred orientation,
+and $\sigma_s$ approximates the radius of the receptive fields.
 
-\begin{figure}
-\centering
-\subfloat[Phase offset $\psi=\pi/2$]{
-  \includegraphics[width=0.4\linewidth]{images/fig4a.jpg}
-  \label{fig:4a}}
-\subfloat[Phase offset $\psi=0$]{
-  \includegraphics[width=0.4\linewidth]{images/fig4b.jpg}
-  \label{fig:4b}}
-\caption{Gabor functions with different phases.}
-\label{fig:4}
-\end{figure}
-
-It is notable that the phase offset $\psi$ affects the pattern of the receptive fields.
-\figurename~\ref{fig:4}~shows two Gabor functions with different phase offsets.
-The sinusoidal factors of the two Gabor functions are odd function and even function respectively.
-The odd Gabor function resembles a gradient operator while the even Gabor function resembles a Laplacian operator.
-In a computational point of view, the Laplacian operator calculates a higher order derivative than the gradient operator.
-The gradient operator suppresses the smooth area in images.
-Moreover, the Laplacian operator is able to suppress area that varies smoothly, which is also neglected by human eyes generally.
-Therefore, in our model, the even Gabor function with the phase offset $\psi=0$ is employed.
-
-The layer of simple cells operates on raw image input. 
+Simple cells operates on raw image input. 
 The output of simple cells with preferred orientation $\theta$ and scale $\sigma_s$ is the following convolution passed through a transfer function $\phi$,
@@
 
-Given $I(x,y)$ as some input image, the output of complex cells
-with the preferred orientation $\theta$ is defined as the following convolution.
+Given $I(x,y)$ as some input image, the output of complex cells with the preferred orientation $\theta$ is defined as the following convolution.
 \begin{equation}
@@
 
-\figurename~\ref{fig:5} shows four image patches.
+\figurename~\ref{fig:4} shows four image patches.
 For each image patch, the output values of complex cells with different preferred orientations (from $0^\circ$ to $180^\circ$) are plotted in a bar chart.
 The charts show the distribution of complex cell activities.
-The entropy calculated accordingly indicates that patches composed of simple structures have non-uniform distributions and thus low entropy values.
+The entropy calculated accordingly indicates that patches composed of geometrically ordered structures have non-uniform distributions and thus low entropy values.
 Therefore, we filter out regions with high entropy.
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig5.jpg}} 
+\centerline{\includegraphics[width=0.8\linewidth]{images/fig4.jpg}} 
 \caption{Entropy of image patches.
@@
 with different preferred orientations (from $0^\circ$ to $180^\circ$).}
-\label{fig:5}
+\label{fig:4}
 \end{figure}
@@
 In addition to entropy, local competition also plays a role in attentional selection \cite{desimone1995}.
-With limited neural resources, only strong and competitive neuronal signals get transmitted and processed.
-In our model, the saliency layer finds local maximums of complex cell output
-and filters out those points with high entropy values or low activities (or output values).
+In our model, the saliency layer finds local maximums of complex cell output and filters out those points with high entropy values or low activities.
 The algorithm is listed in Algorithm \ref{alg:1}.
@@
 \end{algorithm}
+
+\figurename~\ref{fig:5} shows the process of feature detection.
+Saliency filters successfully filter out the trees and handwriting at the background of the images.
+These regions have little shape features but the activity map of complex cells contains high values caused by a large amount of disordered edges. 
+Disordered edges result in high entropy values because the distribution of complex cell activity over different orientations is near uniform.
+Saliency filters filter out these regions by applying a threshold to the entropy map.
+
+\begin{figure}
+\centering
+\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{images/fig5a.jpg}\label{fig:5a}}\hfil
+\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig5c.jpg}\label{fig:5b}}\hfil
+\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig5e.jpg}\label{fig:5c}}\\
+\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{images/fig5b.jpg}\label{fig:5d}}\hfil
+\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig5d.jpg}\label{fig:5e}}\hfil
+\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig5f.jpg}\label{fig:5f}}\\
+\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig5g.jpg}\label{fig:5g}}\hfil
+\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig5h.jpg}\label{fig:5h}}
+\caption{Feature detection. (a)(d) Original images.
+(b)(e) Complex cell activity map. 
+(c)(f) Entropy map.
+In the maps, red color indicates high value and blue color indicates low value.
+(g)(h) Salient points at the scale of 4, 8, and 16 in terms of simple cell radius.}
+\label{fig:5}
+\end{figure}
 
@@
 
-\section{Shape-based feature extraction}\label{sec:4}
+\section{Experiments}\label{sec:4}
 
-In this section, we demonstrate a series of experiments.
-In the experiments, our proposed model extracts local shape features from images.
-Compared with V4 neurons, it shows similar selectivity for different features.
+In this section, we demonstrate several experiments.
+The first two experiments show that the shape selectivity of our model conforms to that of real V4 neurons.
+The following experiments evaluate our model in feature matching and object recognition tasks.
 
-\subsection{Perceptron over Complex Cells}
+\subsection{Perceptron over complex cells}
 
-We have reviewed the shape selectivity of V4 neurons in section \ref{sec:2}.
-In the following experiment, we examined that the output of complex cells provides sufficient information
-for the emergence of neuronal response pattern of V4.
-The experiment involved only the first two layers (i.e., simple cells and complex cells) in our model.
-A perceptron was configured to take complex cells as input.
+With this experiment, we demonstrate that artificial neurons which take input directly from complex cells can imitate neuronal response pattern of V4.
+Therefore, it is feasible that we use a single layer of RBM to extract V4 features from complex cells.
+
+In the experiment, a perceptron was configured to take complex cells as input.
 It was then trained to distinguish between two shapes shown in \figurename~\ref{fig:7a}.
 The shapes were also used to examine the selectivity of V4 in \cite{pasupathy2001}.
-The difference between the two shapes is that one has a sharp projection towards the top right.
+The difference between the two shapes is that one has a sharp corner towards the top right.
 
@@
 \centering
-\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{images/fig-7a.jpg}\label{fig:7a}}\hfil
-\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{images/fig-7b.jpg}\label{fig:7b}}
+\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{images/fig7a.jpg}\label{fig:7a}}\hfil
+\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{images/fig7b.jpg}\label{fig:7b}}
 \caption{Response map and weight matrices of a perceptron that distinguishes two shapes.
@@
 
-Since V4 neurons show a certain degree of invariance in their selectivity across changes in stimulus position,
+Since V4 neurons show a certain degree of invariance in stimulus position,
 we moved the shapes randomly within the receptive field of the perceptron 
@@
 The output of complex cells were passed to the perceptron.
-We had complex cells with different preferred orientations 
-(from $0^\circ$ to $170^\circ$ in steps of $10^\circ$) and over different positions in the receptive field.
+We had complex cells with 18 different preferred orientations (from $0^\circ$ to $170^\circ$ in steps of $10^\circ$) 
+and over different positions in the V4 receptive field.
 Therefore, the input of the perceptron consisted of 18 matrices, 
@@
 
-The trained perceptron exhibited a strong bias towards the shape with convex curvature towards the top right.
+The trained perceptron exhibited a strong bias towards the shape with a sharp corner towards the top right.
 It also showed a significant degree of invariance to the stimulus position.
-It is obvious that the selectivity is not formed from certain excitatory sub-regions
+It is obvious that this kind of selectivity is not formed from certain excitatory sub-regions
 or inhibitory sub-regions of the receptive field which was found of simple cells \cite{hubel1962,pasupathy2001}.
@@
 
-This experiment demonstrates that complex cells provide sufficient information
-for V4 neurons to show selectivity observed in neurobiological experiments.
+This experiment demonstrates that complex cells provide sufficient information for V4 neurons to show selectivity observed in neurobiological experiments.
 
-\subsection{Shape Selectivity}
+\subsection{Shape selectivity}
 
 In the following experiment, we trained our model to learn shapes.
+After training, our model can distinguish the stimuli which V4 neurons are selectively tuned for.
+The training samples were collected from \cite{gallant1996,pasupathy1999,pasupathy2001}, 
+including 4 categories of stimuli, i.e., sinusoidal gratings, non-Cartesian gratings, segmented curves, and closed shapes.
 The training process involved the layers of simple cells, complex cells and RBM encoders.
-The training samples were collected from \cite{gallant1996,pasupathy1999,pasupathy2001}, 
-including 4 categories of stimuli, i.e.,
-sinusoidal gratings, non-Cartesian gratings, segmented curves, and closed shapes.
-The simple cells in our model took these sample images as input.
-The images were processed by simple cells and complex cells.
-Since the sample images were fitted into V4 receptive fields in neurobiological experiments,
-we adjusted the scale of the images to fit the size of a single V4 receptive field in our model too.
-Therefore the saliency filters were not necessary in this experiment.
-The output of complex cells were passed directly on to an RBM encoder.
+Since each sample image was fitted into a single V4 receptive field, the saliency filters were not involved in this experiment.
 
-We used an RBM with 256 hidden units (RBM output neurons) in this experiment.
-In order to assess the selectivity of these units,
-we assigned each unit a selectivity index over a category of stimuli.
-The selectivity index is defined as the ratio of the maximal response to the average response 
-over the stimuli of a certain category.
-Given an RBM output unit $h$ and its output value $h_i$ over a category of $N$ stimuli for $i=1,2,\dots,N$.
+We used an RBM with 256 hidden units in this experiment.
+In order to assess the selectivity of these units, we assigned each unit a selectivity index over a category of stimuli.
+The selectivity index is defined as the ratio of the maximal response to the average response over the stimuli of a certain category.
+Given an RBM output unit $h$ and its output value $h_i$ over the $i$-th stimulus in a category of $N$ stimuli.
 The selectivity index $\mathcal{S}$ of $h$ is given by
@@
 \end{equation}
-In the following statistics, we assumed that a neuron has significant selectivity over a category of stimuli
-if the selectivity index is greater than 3.5.
+In the following statistics, we assumed that a unit has significant selectivity over a category of stimuli
+if its selectivity index is greater than the average index.
 Among the 256 units, 171 units exhibited significant selectivity for different stimuli.
@@
 68 units exhibited selectivity for closed shapes (\figurename~\ref{fig:10}).
-The shape selectivity of our model is listed in \tablename~\ref{tab:2}.
+The shape selectivity of our model is listed in \tablename~\ref{tab:1}.
 The categories of selectivity are not mutually disjoint.
@@
 \toprule
-Selectivity & Number of Units & Percentage \\
+Selectivity & \#Units & Percentage \\
 \midrule
@@
 \end{tabular}
-\label{tab:2}
+\label{tab:1}
 \end{table}
 
-\figurename~\ref{fig:8a} shows the sample patterns of segmented curves.
-\figurename~\ref{fig:8b} to \figurename~\ref{fig:8d} show the response maps of 3 RBM output neurons.
+\figurename~\ref{fig:8} show three response maps to segmented curves.
 Darker background indicates stronger response.
-The selectivity indexes is shown at the top.
-It is shown that the three neurons exhibited strong bias towards curves rather than straight lines.
-The neurons were tuned for the orientation of the projection of the segmented curves.
+It is shown that the three units exhibited strong bias towards curves rather than straight lines.
+They were all tuned for the orientation of the projections.
 This is compliant with the selectivity of V4 neurons described in \cite{pasupathy1999}.
-
-\figurename~\ref{fig:9} shows the response maps of two RBM output neurons
-which are selective for non-Cartesian gratings.
-Both of the two neurons were not sensitive to sinusoidal gratings.
-The one in \figurename~\ref{fig:9a} was selective for a pair of circular sectors.
-The one in \figurename~\ref{fig:9b} preferred non-Cartesian gratings of helix shapes.
-
-\figurename~\ref{fig:10} shows the sample patterns of closed shapes and the response map of two neurons.
-The two neurons were selectively tuned for two kinds of shapes.
-They were sensitive to the angular position of the shapes.
+\figurename~\ref{fig:9} shows the response maps of two units which are selective for non-Cartesian gratings.
+Both of the two units were not sensitive to sinusoidal gratings.
+\figurename~\ref{fig:10} shows the sample patterns of closed shapes and the response map of two units selectively tuned for two kinds of shapes.
+They were sensitive to the angular position of the shapes \cite{pasupathy2001}.
 
@@
 \centering
-\subfloat[Patterns]{\includegraphics[width=0.24\linewidth]{images/fig-8a.jpg}\label{fig:8a}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8b.jpg}\label{fig:8b}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8c.jpg}\label{fig:8c}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.24\linewidth]{images/fig-8d.jpg}\label{fig:8d}}
-\caption{RBM output neurons responding to convex curves towards certain directions.
+\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8a.jpg}\label{fig:8a}}\hfil
+\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8b.jpg}\label{fig:8b}}\hfil
+\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8c.jpg}\label{fig:8c}}\hfil
+\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8d.jpg}\label{fig:8d}}
+\caption{RBM output units responding to convex curves towards certain directions.
 (a) Sample patterns of segmented curves.
-(b) Response map of a neuron tuned for curves projecting upwards.
-(c) Response map of a neuron tuned for curves projecting downwards.
-(d) Response map of a neuron tuned for curves projecting towards the top left.}
+(b-d) Selectivity indexes and response maps of units tuned for upward, downward, and top-left projections. }
 \label{fig:8}
@@
 \centering
-\subfloat[Response map of a neuron selective for a pair of circular sectors (selectivity index = 3.92).]%
-{\includegraphics[width=0.9\linewidth]{images/fig-9a.jpg}\label{fig:9a}}\\
-\subfloat[Response map of a neuron selective for helix shapes (Selectivity index = 6.37).]%
-{\includegraphics[width=0.9\linewidth]{images/fig-9b.jpg}\label{fig:9b}}
-\caption{Response maps of RBM output neurons tuned for non-Cartesian gratings.}
+\subfloat[Response map of a unit selective for a pair of circular sectors (selectivity index = 3.92).]%
+{\includegraphics[width=0.9\linewidth]{images/fig9a.jpg}\label{fig:9a}}\\
+\subfloat[Response map of a unit selective for helix shapes (Selectivity index = 6.37).]%
+{\includegraphics[width=0.9\linewidth]{images/fig9b.jpg}\label{fig:9b}}
+\caption{Response maps of RBM units tuned for non-Cartesian gratings.}
 \label{fig:9}
@@
 \centering
-\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{images/fig-10a.jpg}\label{fig:10a}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10b.jpg}\label{fig:10b}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig-10c.jpg}\label{fig:10c}}
-\caption{RBM output neurons responding to closed shapes.}
+\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{images/fig10a.jpg}\label{fig:10a}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig10b.jpg}\label{fig:10b}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig10c.jpg}\label{fig:10c}}
+\caption{RBM units responding to closed shapes.}
 \label{fig:10}
@@
 
-This experiment demonstrates that our model can learn to distinguish the stimuli
-which V4 neurons are selectively tuned for.
-
-\subsection{Feature detection}
-
-In the above experiments, it has been demonstrated that a unit of the RBM hidden layer 
-can encode the shape feature in the receptive field of a single V4 computation unit.
-When we deal with larger images, especially natural images,
-a full scan over the whole image is not computationally economic.
-It also complicates succeeding processing for high-level tasks such as object recognition and scene understanding.
-We have to find some regions of interest 
-and focus the computation units over a limited number of interesting regions.
-The visual neural system takes a similar approach.
-As we have reviewed in previous sections, V4 is closely related with selective visual attention \cite{roe2012}.
-Research shows that V4 receptive field shrinks and shifts towards saccade target \cite{hamker2006}.
-We simulated such visual attention with the saliency filters in our model.
-In the following experiment,
-we evaluated the saliency filters with feature detection experiment.
-
-\begin{figure}
-\centering
-\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{images/fig-11a.jpg}\label{fig:11a}}\hfil
-\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11c.jpg}\label{fig:11b}}\hfil
-\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11e.jpg}\label{fig:11c}}\\
-\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{images/fig-11b.jpg}\label{fig:11d}}\hfil
-\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig-11d.jpg}\label{fig:11e}}\hfil
-\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig-11f.jpg}\label{fig:11f}}\\
-\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11g.jpg}\label{fig:11g}}\hfil
-\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig-11h.jpg}\label{fig:11h}}
-\caption{Feature detection. (a) and (d) show the original images.
-(b) and (e) are the complex cell activity. 
-Red color indicates strong activity and blue color indicates inhibition.
-(c) and (f) are the entropy map.
-Red color indicates high entropy value and blue color indicates low entropy value.
-(g) and (h) are the selected salient points
-at the scale of 4, 8, and 16 in terms of simple cell radius.}
-\label{fig:11}
-\end{figure}
-
-\figurename~\ref{fig:11} shows the process of feature detection with the saliency filters in our model.
-\figurename~\ref{fig:11b} and \figurename~\ref{fig:11d} show the activity map of complex cells.
-The activities are high in the region of the trees behind the car (\figurename~\ref{fig:11a})
-and the handwriting on the white board (\figurename~\ref{fig:11c}).
-Saliency filters filter out these regions because these regions have a comparatively high entropy.
-These areas consist of a large amount of disordered edges
-which result in a near uniform distribution of complex cell activity over preferred orientations
-and thus high entropy values.
-The saliency filters can find salient points at each given scale.
-\figurename~\ref{fig:11g} and \figurename~\ref{fig:11h} show the salient points
-at the scale of 4, 8, and 16 in terms of the simple cell radius
-(The size of the other units can be inferred with the relationship
-shown in Table \ref{tab:1}).
-
-\section{Object discrimination}
-\label{sec:5}
-
-In this section, we demonstrate the performance of the features 
-generated by our model in feature matching and object recognition tasks.
-
-We do not use typical criteria for evaluating local descriptors \cite{mikolajczyk2005}
-for the reason that the typical criteria focus on
-matching the same physical point in different viewing conditions,
-at which humans or other intelligent species
-are not significantly better than computer systems.
-The human vision excels in image understanding tasks
-such as recognizing a new instance of some category
-with only few prototypes for learning.
-Our model conforms to the neural mechanism of vision and
-thus exhibits similar learning ability in the experiments.
-
 \subsection{Feature matching}
 
-Traditional local features excel in matching features
-corresponding to the same physical point on an object
-seen from different viewpoints \cite{mikolajczyk2005}.
-These features are usually very selective for a target object
-but lack invariance with respect to different objects from a target category.
-However, the generalization ability within object categories
-is important for image understanding tasks such as object recognition.
+In this section, we demonstrate the performance of the features generated by our model in feature matching tasks.
+We do not use typical criteria for evaluating local descriptors \cite{mikolajczyk2005}
+for the reason that the typical criteria focus on matching the same physical point in different viewing conditions,
+at which humans or other intelligent species are not significantly better than computer systems.
+The human vision excels in image understanding tasks such as recognizing a new instance from some category.
+It requires generalization ability of the features.
 
-We provide a qualitative evaluation of our features
-by matching features between different images.
+We first provide a qualitative evaluation.
 The SIFT features are used for comparison.
-The VLFeat open-source library \cite{vedaldi2008}
-were utilized for the SIFT implementation 
-and the feature matching algorithm.
+The VLFeat open-source library \cite{vedaldi2010} were utilized for the SIFT implementation and the feature matching algorithm.
 
@@
   \subfloat[Matching between the same face]{
-    \includegraphics[width=0.8\linewidth]{images/fig-12a.jpg}
-    \label{fig:12a}}\\
+    \includegraphics[width=0.8\linewidth]{images/fig11a.jpg}
+    \label{fig:11a}}\\
   \subfloat[Matching between different faces]{
-    \includegraphics[width=0.8\linewidth]{images/fig-12b.jpg}
-    \label{fig:12b}}
+    \includegraphics[width=0.8\linewidth]{images/fig11b.jpg}
+    \label{fig:11b}}
 \caption{Matching face images.}
-\label{fig:12}
+\label{fig:11}
 \end{figure}
 
-\figurename~\ref{fig:12} shows matchings of features in face images.
-As shown in \figurename~\ref{fig:12a},
-SIFT features match perfectly between face images of the same person.
-The neural features generated by our model achieve competitive distinctiveness.
-However, \figurename~\ref{fig:12b} shows that
-matching of SIFT features between face images of different people
-is quite disordered.
+\figurename~\ref{fig:11} shows matchings of features in face images.
+As shown in \figurename~\ref{fig:11a}, both SIFT features and neural features match perfectly between face images of the same person.
+However, \figurename~\ref{fig:11b} shows that matching of SIFT features between face images of different people is quite disordered.
 SIFT features excel in matching highly distinctive features under image transformations
-but lack the generalization ability 
-to capture variations in objects appearance of the same category.
-By contrast, the neural features show robust performance 
-in matching features from different faces.
-The result suggests that our model exhibits 
-both competitive discrimination ability
-and better generalization ability compared with SIFT features.
+but lack the generalization ability to capture variations in objects appearance of the same category.
+By contrast, the neural features show robust performance in matching features from different faces.
+The result suggests that our model exhibits both competitive discrimination ability and better generalization ability compared with SIFT features.
 
@@
   \subfloat[Our match]{
-    \includegraphics[width=0.8\linewidth]{images/fig-13a.jpg}
-    \label{fig:13a}}\\
+    \includegraphics[width=0.8\linewidth]{images/fig12a.jpg}
+    \label{fig:12a}}\\
   \subfloat[SIFT match]{
-    \includegraphics[width=0.8\linewidth]{images/fig-13b.jpg}
-    \label{fig:13b}}
+    \includegraphics[width=0.8\linewidth]{images/fig12b.jpg}
+    \label{fig:12b}}
 \caption{Comparison of feature matching between van images.}
+\label{fig:12}
+\end{figure}
+
+Another comparison in \figurename~\ref{fig:12} demonstrates
+more clearly the advantage of our model compared with SIFT feature.
+\figurename~\ref{fig:12a} shows that our model produced correct matches
+between the model van in the left and the two real vans in the right. 
+However, SIFT feature failed to produce correct matches (\figurename~\ref{fig:12b}). 
+SIFT feature utilizes the direction of image gradient. 
+It fails to capture the shape feature of an object when the object's color changes dramatically because the direction of image gradient changes with the color. 
+Our model focuses on the shape feature and thus achieves a stable matching despite the change in color.
+
+For quantitative evaluation, we use the ``Face-easy'' category from the Caltech 101 dataset \cite{fei2007}.
+This category consists of frontal face photos of many different people.
+They are all cropped to the same size and the same face location.
+The images are randomly paired for feature matching.
+A correct match between facial features should be corresponding regions in the two images.
+We use precision-recall curves to compare our neural features with the SIFT features.
+\begin{equation}
+\text{Precision} = \frac{\text{\#correct matches}}{\text{\#correct matches} + \text{\#false matches}}.
+\end{equation}
+\begin{equation}
+\text{Recall} = \frac{\text{\#correct matches}}{\text{\#all features}}.
+\end{equation}
+The result is shown in \figurename~\ref{fig:13}.
+Extensive evaluation on the generalization ability of our features is in the object recognition experiment in the next subsection.
+
+\begin{figure}
+\centerline{\includegraphics[width=0.9\linewidth]{images/fig13.jpg}} 
+\caption{Matching facial features in Caltech-101 ``Face-easy'' category.}
 \label{fig:13}
@@
 
-Another comparison in \figurename~\ref{fig:13} demonstrates
-more clearly the advantage of our model compared with SIFT feature.
-\figurename~\ref{fig:13a} shows that our model produced correct matches
-between the model van in the left and the two real vans in the right. 
-However, SIFT feature failed to produce correct matches (\figurename~\ref{fig:13b}). 
-SIFT feature utilizes the direction of image gradient. 
-It fails to capture the shape feature of an object when
-the object's color changes dramatically because the direction of
-image gradient changes with the color. 
-Our model focuses on the shape feature and thus achieves a stable matching despite
-the change in color.
-
-\subsection{Object recognition layer}
-
-As we have reviewed in section 2,
-the IT cortex is where object recognition is finally achieved in the brain.
-The exact nature of the high-level visual features 
-encoded by IT neurons remains to be determined.
-Yet research shows that V4 neurons encode shapes by population responses \cite{pasupathy2002}.
-It is likely that an IT neuron collects the responses from a group of afferent V4 neurons 
-and synthesizes a unified representation of shapes.
-In our model, the object recognition layer consists of IT computation units.
-Each IT unit combines the output of several V4 units into a local image feature.
-The receptive fields of these V4 units scatter around the center of the IT receptive field.
-Though there is no anatomical evidence showing that 
-the spatial distribution of the V4 receptive fields 
-is necessarily restricted to any fixed patterns,
-in our model, we choose afferent V4 units from a group of rectangular grid points.
-Each IT unit has 9 afferent connections from the V4 layer.
-The 9 connected V4 units are located in a $3\times 3$ grid.
-\figurename~\ref{fig:13} demonstrates an IT receptive field 
-and the receptive fields of the connected V4 neurons.
-The demonstration is shown over
-the output of complex cells given the image of Lena as input.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{images/fig-14.jpg}
-\caption{Demonstration of IT unit in our model. 
-IT unit combines V4 inputs in a grid layout rotated to the strongest orientation.}
-\label{fig:14}
-\end{figure}
-
-In order to achieve the invariance of feature orientations,
-the gird is rotated to the strongest orientation at the center of the receptive field.
-We have been inspired by the SIFT \cite{lowe1999} descriptor.
-The rotation is also demonstrated in \figurename~\ref{fig:14}.
-The grid is rotated to the orientation of the edge of the hat appearing at the center of the grid.
-The strongest central orientation can be obtained by taking a maximal value 
-over the output of the complex cells connected to the central V4 units.
-
-The output of an IT unit is thus defined as a concatenation of the V4 feature vectors.
-\begin{equation}
-\mathbf{F}(x,y) = \mathbf{V}(\hat{x}_1,\hat{y}_1)\|
-\mathbf{V}(\hat{x}_2,\hat{y}_2)\|
-\mathbf{V}(\hat{x}_3,\hat{y}_3)\|{}  \cdots\|
-\mathbf{V}(\hat{x}_n,\hat{y}_n),
-\end{equation}
-where $(\hat{x}_i,\hat{y}_i)^{\top} = 
-\left(\begin{array}{ccc}
-\cos\hat\theta & -\sin\hat\theta & x \\
-\sin\hat\theta & \cos\hat\theta & y
-\end{array}\right)\cdot(x_i,y_i,1)^{\top}$ for $i=1,2,3,\dots,n$ and 
-$\hat\theta = \arg\!\max_\theta C_\theta(x,y)$.
-In this equation, $\mathbf{F}$ represents the image feature generated by IT units.
-$\hat\theta$ represents the strongest orientation at the center of the receptive field,
-which is calculated over the output of complex cells $C_\theta$.
-$(\hat{x}_i,\hat{y}_i)$ is a group of rotated grid points
-indicating the centers of receptive fields of afferent V4 units.
-$(x_i,y_i)$ is a group of grid points.
-In our model, $(x_i,y_i)$ is defined as the $3\times 3$ grid points fitted 
-in a rectangular region.
-
 \subsection{Object recognition}
 
-We evaluate the performance of the proposed model
-in object recognition in clutter for which 
-the target object in both the training and test samples
-appears at variable scales and positions 
-within an unsegmented image.
+We evaluate the performance of the proposed model in object recognition in clutter for which 
+the target object in both the training and test samples appears at variable scales and positions within an unsegmented image.
+Features at the salient-points are extracted and passed to a classifier.
+We experimented with the SVM classifier and evaluated the classifier over the Caltech 101 dataset \cite{fei2007}.
 
-The bags of key-points approach \cite{csurka2004} are employed.
-Key-points are detected with the saliency sublayer of the V4 computation units in our model.
-Features at the key-points are extracted with IT units and passed to a classifier.
-We experimented with the SVM classifier
-and evaluated the classifier over the Caltech 101 dataset \cite{feifei2004}.
-
-The parameters that controls the size of simple, complex and V4 receptive field
-are summarized in \tablename~\ref{tab:1} in section 3.
-In addition to these parameters,
-the sampling density of different preferred orientations is also a parameter in our model.
-In the RBM encoder sublayer of V4 units,
-the visible RBM units collect their input from complex cells 
-with different preferred orientations.
-A sampling density of $n$ means that a preferred orientation is taken from the range
-$\{0^{\circ}, 180^{\circ}/n, 2\times 180^{\circ}/n, \dots, (n-1)\times 180^{\circ}/n\}$.
-We investigate the selection of sampling density of preferred orientations.
-30 training samples are randomly chosen from each category.
+In the RBM encoders of V4 units, the visible RBM units collect their input from complex cells with different preferred orientations.
+Therefore, we investigate the selection of sampling density of orientations (i.e., the number of different preferred orientations).
+The result is shown in \figurename~\ref{fig:14a}.
+In this experiment, 30 training samples are randomly chosen from each category.
 For each training image, 50 features are randomly chosen and passed to the classifier.
-The results reported are averaged over 8 independent runs.
+The result shows that selection of sampling density does not exhibit obvious impact on the performance.
+With a small density, we can lower the dimension of the input of RBM encoders without a significant loss in object recognition performance.
+Thus in the succeeding experiments, a sampling density of $3$ is used.
 
@@
   \subfloat[Sampling density of orientations]{
-    \includegraphics[width=0.9\linewidth]{images/fig-15a.eps}
-    \label{fig:15a}}\\
+    \includegraphics[width=0.9\linewidth]{images/fig14a.eps}
+    \label{fig:14a}}\\
   \subfloat[Number of features per image]{
-    \includegraphics[width=0.9\linewidth]{images/fig-15b.eps}
-    \label{fig:15b}}
+    \includegraphics[width=0.9\linewidth]{images/fig14b.eps}
+    \label{fig:14b}}
 \caption{Performance with different parameters.}
-\label{fig:15}
+\label{fig:14}
 \end{figure}
 
-As shown in \figurename~\ref{fig:15a},
-the selection of sampling density does not exhibit obvious impact on the performance.
-With a small density, we can lower the dimension of the input of RBM encoders
-without a significant loss in object recognition performance.
-Thus in the succeeding experiments, a sampling density of $3$ is used.
+We can also control the number of features extracted from an image by adjusting the threshold in the saliency filters. 
+To investigate the contribution of the number of features on performance, we vary the number of features chosen from each training image.
+We choose 20 training samples from each category.
+The result is shown in \figurename~\ref{fig:14b}.
+With 50 features per image the result approaches the best performance.
+In the succeeding experiments, we take 50 features from each image.
 
-We can also manipulate the saliency threshold in the saliency sublayer of V4 units
-to change the number of key-points detected in an image,
-which inturn changes the number of features generated by IT units.
-To investigate the contribution of the number of features on performance,
-we vary the number of features chosen from each training image.
-We choose 20 training samples from each category.
-The results reported are also averaged over 8 independent runs.
-\figurename~\ref{fig:15b} shows that 
-with 50 features per image the result approaches the best performance.
-The number of required features can be notably reduced by introducing 
-a clustering step in the bag of words approach \cite{csurka2004}.
-However, Serre et al. \cite{serre2007} reported that clustering could
-be sensitive to the choice of parameters and initializations,
-leading poorer results.
-
-The influence of the number of training samples 
-on the performance of our model is also studied.
-For each category, we randomly select training sets of size
-5,10,15,20,25 and 30.
-\figurename~\ref{fig:16a} shows the performance as
-a function of the number of training images.
-The results reported are averaged over all categories
-with 50 features extracted from each training image.
-The performance is compared with other published results.
+The influence of the number of training samples on the performance of our model is also studied.
+For each category, we randomly select training sets of size 5,10,15,20,25 and 30.
+\figurename~\ref{fig:15a} shows the performance as a function of the number of training images.
+The results reported are averaged over all categories.
+The performance is compared with other published results \cite{grauman2005,lazebnik2006,zhang2006,wang2006,bosch2007,boiman2008,liu2013,heo2014}.
 
@@
   \subfloat[Comparison with other methods]{
-    \includegraphics[width=0.9\linewidth]{images/fig-16a.eps}
-    \label{fig:16a}}\\
+    \includegraphics[width=0.9\linewidth]{images/fig15a.eps}
+    \label{fig:15a}}\\
   \subfloat[Comparison with other features]{
-    \includegraphics[width=0.9\linewidth]{images/fig-16b.eps}
-    \label{fig:16b}}
+    \includegraphics[width=0.9\linewidth]{images/fig15b.eps}
+    \label{fig:15b}}
 \caption{Comparison of performance.}
-\label{fig:16}
+\label{fig:15}
 \end{figure}
 
-\figurename~\ref{fig:16b} shows
-the performance of our model over the 101 categories
-compared with SIFT features \cite{lowe1999},
-PHOW features \cite{lazebnik2006},
-and biologically inspired HMAX model \cite{serre2007}.
-The results are obtained with 30 training images randomly selected 
-from each category.
-SIFT features and neural features generated with our model
-are passed to an SVM classifier.
-The images passed to the HMAX model are resized 
-to a uniform height of 140 pixels as recommended by \cite{serre2007}.
-Our model exhibits competitive performance 
-compared with other models.
-As shown in \figurename~\ref{fig:16b},
-the SIFT performance points over the 101 categories
-are almost completely to the left of the equivalent line.
-The PHOW performance points spread through 
-a wider range from $0.5$ to $1$ indicating 
-the lack of robustness compared with our model.
-Compared with the biologically inspired HMAX model,
-our performance is also better for more than half
-of the 101 categories.
+\figurename~\ref{fig:15b} shows the performance of our model over the 101 categories
+compared with SIFT features \cite{lowe1999}, PHOW features \cite{lazebnik2006}, and biologically inspired HMAX model \cite{cadieu2007}.
+The results are obtained with 30 training images randomly selected from each category.
+Our model exhibits competitive performance compared with other models.
+As shown in \figurename~\ref{fig:15b}, the SIFT performance points over the 101 categories are almost completely to the left of the equivalent line.
+The PHOW performance points spread through a wider range from $0.5$ to $1$ indicating the lack of robustness compared with our model.
+Compared with the biologically inspired HMAX model, our performance is also better for more than half of the 101 categories.
 
-\section{Conclusion}\label{sec:6}
+\section{Discussion and conclusion}\label{sec:5}
 

### 2015-09-06 article.tex
finish revision

@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centerline{\includegraphics[width=0.8\linewidth]{images/fig1.jpg}} 
@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centering
@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centerline{\includegraphics[width=\linewidth]{images/fig3.eps}} 
@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centerline{\includegraphics[width=0.8\linewidth]{images/fig4.jpg}} 
@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centerline{\includegraphics[width=0.99\linewidth]{images/fig6.eps}} 
@@
 \centering
-\begin{tabular}{rrr}
-\toprule
-Selectivity & \#Units & Percentage \\
-\midrule
-Sinusoidal gratings & 50 & 19.5\% \\
-Non-Cartesian gratings & 55 & 21.5\% \\
-Segmented curves & 102 & 39.8\% \\
-Closed shapes & 68 & 26.6\% \\
-\bottomrule
+\begin{tabular}{|r|r|r|}
+\hline
+Selectivity & \#Units & Percentage \\\hline
+Sinusoidal gratings & 50 & 19.5\% \\\hline
+Non-Cartesian gratings & 55 & 21.5\% \\\hline
+Segmented curves & 102 & 39.8\% \\\hline
+Closed shapes & 68 & 26.6\% \\\hline
 \end{tabular}
@@
 We first provide a qualitative evaluation.
-The SIFT features are used for comparison.
+The SIFT features \cite{lowe1999} are used for comparison.
 The VLFeat open-source library \cite{vedaldi2010} were utilized for the SIFT implementation and the feature matching algorithm.
 
-\begin{figure}
+\begin{figure}[!t]
 \centering
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=0.9\linewidth]{images/fig13.jpg}} 
+\begin{figure}[!t]
+\centerline{\includegraphics[width=0.9\linewidth]{images/fig13.eps}} 
 \caption{Matching facial features in Caltech-101 ``Face-easy'' category.}
@@
 
-\subsection{Traditional image features}
+\begin{table*}[!t]
+\caption{Comparison of different models}
+\centering
+\small
+\begin{tabular}{|p{0.15\linewidth}|p{0.22\linewidth}|p{0.22\linewidth}|p{0.22\linewidth}|}
+\hline
+& Our model & HMAX & Deep learning \\\hline
+Number of layers & 3 & 2 & $2\sim8$ \\
+& ($+$ saliency filters) & ($+2$ pooling layers) & ($+2\sim3$ pooling layers)\\\hline
+Input images & various size & various size & fixed size \\
+& (multi-scale salient point detector) & (full-scan with fixed-size window) & \\\hline
+Training samples & $5\sim50$ & $5\sim50$ & $>10^4$ \\
+& (per category) & (per category) & (for unsupervised pre-training) \\\hline
+Training & restricted Boltzmann & greedy search for fitting & restricted Boltzmann \\
+algorithms & machine (only V4 layer) & V4 neuron; radial basis function network for image feature extraction & machine (all layers) \\\hline
+\end{tabular}
+\label{tab:2}
+\end{table*}
 
-The model proposed in this paper is a biologically motivated model for image feature extraction.
-As reviewed in \cite{mikolajczyk2005}, traditional image features fall into three categories: 
-distribution-based descriptors, spatial-frequency descriptors and differential-based descriptors.
-Distribution-based descriptors use histograms to represent characteristics of local image regions.
-SIFT descriptor \cite{lowe1999} and shape context \cite{belongie2002} are typical examples of such descriptors.
-Spatial-frequency descriptors and differential-based descriptors use filters and derivatives to capture the invariant features. 
-Examples of these categories are complex filters \cite{schaffalitzky2002} and Baumberg's affine invariant descriptor \cite{baumberg2000}.
-These traditional image features utilize flat models which compute features directly from the image via statistical or algebraic techniques. 
-They have little in common with the neural systems.
+The model proposed in this paper is a biologically motivated model for shape-based image feature extraction.
+Traditional image descriptors utilize flat models which extract features directly from the image via statistical or algebraic techniques \cite{mikolajczyk2005}. 
 On the contrary, our proposed model simulates the biological vision in the level of neural mechanism.
-It is a multilayer model which employs the hierarchical structure of neural systems and extracts information from simple features to gradually complex features.
+Its multilayer structure resembles the neural systems and extracts information from simple features to gradually more complex features.
 
-This process requires a large number of training samples (\cite{krizhevsky2012} uses 1.2 million training images)
-and a neural network of very large scale (In \cite{krizhevsky2012}, the network has 8 layers and more than 400,000 neurons).
-Deep learning can be viewed as automatic feature extraction.
-In our proposed model, we combine automatic feature extraction with artificial feature extraction.
+Previous work on biologically inspired models shows that efficient object recognition algorithms can be developed 
+based on the representation generated by the primary visual cortex \cite{wei2014,wei2014b,zhao2013}.
+In this paper, we investigate deeper into higher levels of the visual cortex.
+We simulate the shape selectivity in area V4 because the shape or contour of an object is generally stable and invariant, which provides a good basis for recognition.
+
+The HMAX model \cite{cadieu2007} and the deep learning model \cite{bengio2009}, which both draw inspiration from the neural systems, 
+construct similar multilayer structures.
+In particular, the deep model exhibits high performance in classifying very large image database \cite{krizhevsky2012}.
+However, a deep model is a neural network of very large scale (In \cite{krizhevsky2012}, the network has 8 layers and more than 400,000 neurons),
+and training such a model generally requires a large number of training samples (\cite{krizhevsky2012} uses 1.2 million training images).
+Deep learning is a process of fully automatic feature extraction.
+Limited training samples can not provide enough variety to keep the model from over-fitting.
+The HMAX model consists of fewer layers and only one layer (S2 layer) is trainable, which allows the model to work on limited train samples.
+In our proposed model, we combine these two approach.
 With prior knowledge of the cortical neurons, we manually design the layers corresponding to the primary visual cortex
 and V4 features are automatically extracted based on the output of the artificially designed network.
-This strategy saves us from retraining the first layers every time
-and it allows the model to work on limited number of training images.
+This strategy saves us from retraining the first layers every time.
+Additionally, we implement a layer of saliency filters to detect key-points in images,
+which enables our model to process images of various sizes efficiently. 
+In \tablename~\ref{tab:2}, we summarize the comparison of the three models.
 
-In this paper, we propose a novel model for the visual area V4.
-It is based on the neural mechanism of the ventral visual pathway.
-We focus on visual attention and shape selectivity in V4.
-V4 in our model is implemented as a multilayer neural network 
-which selects salient points and encodes the shape feature in the neighborhood of the salient points.
-We demonstrate a variety of experiments in which the model was evaluated.
-The results show that the proposed model is consistent with the shape selectivity of area V4.
-It can find salient points in images and encode local shape feature into a discriminative representation.
-
-We further evaluate our model in object recognition experiment.
-We designed a image descriptor which synthesizes several V4 computation units
-into an IT unit to produce a set of invariant local image features.
-Compared with traditional image descriptors,
-the proposed model shows competitive performance in object recognition experiment. 
-It captures the essential shape feature of objects
-and thus achieves a good balance between discrimination and generalization.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{images/fig-17.jpg}
-\caption{Framework of the proposed model.}
-\label{fig:17}
-\end{figure}
-
-The framework of the proposed multilayer neural network model is shown in \figurename~\ref{fig:17}.
-Our contribution is focused on the V4 layer,
-where salient shape features are detected and located,
-and the features are encoded and further synthesized to produce discriminative representation of object shapes.
-The proposed model demonstrates a successful attempt to construct model of the ventral visual pathway
-in the brain for visual discrimination.
-We explicitly adopt the model of several layers of this pathway,
-including well studied lower-level layers, such as simple cells and complex cells.
-We focus our work on V4, which is significant but lacks mature computation models.
-
-Future work should involve improvement of the IT unit and further evaluation on learning and detection shapes.
-The implementation of the IT unit should conform to neural mechanism
-and should be capable of learning generalized representation of shape categories.
-Information about color and texture of objects is also very important,
-which should also participate in the process of object discrimination.
+Future work should involve further investigation into the visual systems and extensive evaluation on shape-based object detection.
+Currently the model extracts only shape-based features.
+Information about color and texture is also very important, which should also participate in the process of visual recognition.
+The proposed model is a feed-forward network.
+We should also consider the function of feedback connections in the neural systems.
 These show the direction of our future work.

### 2015-09-07 article.tex
refine some figures

@@
 Limited training samples can not provide enough variety to keep the model from over-fitting.
-The HMAX model consists of fewer layers and only one layer (S2 layer) is trainable, which allows the model to work on limited train samples.
-In our proposed model, we combine these two approach.
-With prior knowledge of the cortical neurons, we manually design the layers corresponding to the primary visual cortex
-and V4 features are automatically extracted based on the output of the artificially designed network.
+The HMAX model consists of fewer layers and only one layer (S2 layer) is trainable, which allows the model to work on limited training samples.
+In the proposed model, we combine these two approaches.
+The layers corresponding to the primary visual cortex is artificially designed with prior knowledge of the cortical neurons.
+V4 features are automatically extracted based on the output of the artificially designed network.
 This strategy saves us from retraining the first layers every time.
 Additionally, we implement a layer of saliency filters to detect key-points in images,
-which enables our model to process images of various sizes efficiently. 
+which enables the model to process images of various sizes efficiently. 
 In \tablename~\ref{tab:2}, we summarize the comparison of the three models.
@@
 Future work should involve further investigation into the visual systems and extensive evaluation on shape-based object detection.
-Currently the model extracts only shape-based features.
-Information about color and texture is also very important, which should also participate in the process of visual recognition.
-The proposed model is a feed-forward network.
-We should also consider the function of feedback connections in the neural systems.
-These show the direction of our future work.
 

### 2015-09-07 article.tex
fix typo

@@
 In this paper, we demonstrate a novel model of visual area V4 and its application to shape-based feature extraction and object discrimination.
-Experiment results show that the proposed model exhibits similar shape selectivity to that of real V4 neurons
+Experimental results show that the proposed model exhibits similar shape selectivity to that of real V4 neurons
 and it outperforms many existing methods in computer vision tasks including feature matching and object recognition.
@@
 Moreover, this multilayer neural network shows a successful attempt to integrate different stages of visual processing in the brain. 
-It provides an extensible framework so that future findings in cortex mechanisms can be adopted and applied to computer vision tasks.
+It provides an extensible framework so that future findings in cortical mechanisms can be adopted and applied to computer vision tasks.
 
-We briefly introduce some related work on image features and V4 models in section~\ref{sec:2}.
-In section~\ref{sec:3}, we give the detailed description of our proposed model.
+We briefly introduce some related work on V4 models in section~\ref{sec:2}.
+In section~\ref{sec:3}, we give the detailed description of the proposed model.
 It is capable of locating salient points and generating discriminative local shape features around salient points.
-In section~\ref{sec:4}, we show several experiments in which we evaluate the proposed model.
+In section~\ref{sec:4}, we show several experiments in which we evaluate the model.
 The model is compared with real V4 neurons in a shape distinction experiment.
@@
 
-Complex cells are modeled as Gaussian filters which take inputs from simple cells,
+Complex cells are modeled as Gaussian filters which take input from simple cells,
 \begin{equation}

### 2015-09-08 article.tex
finish revision

@@
 
+The rest of the paper is organized as follows.
 We briefly introduce some related work on V4 models in section~\ref{sec:2}.
@@
 
-In this section, we briefly review some existing models of V4 and other related models.
+In this section, we briefly review some existing models of V4 and the deep learning model.
 
@@
 Several models have been proposed for V4 neurons.
-We briefly introduce the SRF model \cite{david2006} and the HMAX model \cite{cadieu2007}.
+We focus on two popular models, i.e., the SRF model \cite{david2006} and the HMAX model \cite{cadieu2007}.
 The spectral receptive field (SRF) \cite{david2006} describes properties of V4 receptive field in terms of the orientation and spatial frequency spectrum.
@@
 Complex cells have larger receptive fields.
-A stimulus is effective wherever it is placed in the complex receptive field, provided that the orientation is appropriate \cite{hubel1962}.
+A stimulus is effective wherever it is placed in the receptive field, provided that the orientation is appropriate \cite{hubel1962}.
 
@@
 
-Simple cells operates on raw image input. 
+Simple cells operate on raw image input. 
 The output of simple cells with preferred orientation $\theta$ and scale $\sigma_s$ is the following convolution passed through a transfer function $\phi$,
@@
 The radius of complex receptive field is approximately twice as long as that of a simple cell.
-We take this size because qualitatively the complex receptive fields are usually two to five times wider 
+We take this size because the complex receptive fields are usually two to five times wider 
 than the bars of gratings that stimulate them most effectively \cite{hubel1962}.
@@
 The operation can be viewed as a Gaussian-weighted average pooling.
-We choose weighted average pooling because average pooling can be efficiently calculated with artificial neurons
-and weighted average gives the complex cell's receptive field a smooth and circular boundary.
+We choose this pooling method because 
+(1) the Gaussian weight function gives the pooling region (i.e., the receptive field of complex cells) a smooth and circular boundary
+and (2) average pooling can be efficiently calculated with artificial neurons.
 
@@
 The underlying neural mechanism is unknown.
-V4 neurons exhibit preference for geometrically ordered stimuli \cite{sasaki2005}.
+However, research suggests that V4 neurons exhibit preference for geometrically ordered stimuli \cite{sasaki2005}.
 Therefore we use entropy (a measure of disorder) to measure the saliency of images.
 The entropy is calculated according to the output of complex cells.
-Given a point $(x,y)$, in the neighborhood of $(x,y)$,
-a complex cell with preferred orientation $\theta$ has output value $C_{\theta}(x,y)$,
+In the neighborhood of a given point $(x,y)$,
+a complex cell with the preferred orientation $\theta$ has output value $C_{\theta}(x,y)$,
 We suppose that the probability of a complex cell being activated is proportional to the output value.
@@
 The charts show the distribution of complex cell activities.
-The entropy calculated accordingly indicates that patches composed of geometrically ordered structures have non-uniform distributions and thus low entropy values.
-Therefore, we filter out regions with high entropy.
+Patches composed of geometrically ordered structures have non-uniform distributions, which result in low entropy values.
+These patches contain salient shape features which should be extracted.
+Conversely, patches with high entropy values should be filtered out.
 
@@
 
-\figurename~\ref{fig:5} shows the process of feature detection.
+\figurename~\ref{fig:5} demonstrates the process of detecting salient points.
 Saliency filters successfully filter out the trees and handwriting at the background of the images.
-These regions have little shape features but the activity map of complex cells contains high values caused by a large amount of disordered edges. 
-Disordered edges result in high entropy values because the distribution of complex cell activity over different orientations is near uniform.
+In these regions, disordered edges result in high entropy values because the distribution of complex cell activity over different orientations is near uniform.
 Saliency filters filter out these regions by applying a threshold to the entropy map.
@@
 \subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig5h.jpg}\label{fig:5h}}
-\caption{Feature detection. (a)(d) Original images.
+\caption{Detecting salient points. (a)(d) Original images.
 (b)(e) Complex cell activity map. 
@@
 
-We followed \cite{hinton2010} for the training algorithm of RBM and the choice of parameters.
+We follow \cite{hinton2010} for the training algorithm of RBM and the choice of parameters.
 Weight-decay is used to reduce overfitting by adding an L2 penalty term, $\frac{1}{2}\lambda w^2$, to the energy function.
@@
 
-In this section, we demonstrate several experiments.
+In this section, we demonstrate four experiments.
 The first two experiments show that the shape selectivity of our model conforms to that of real V4 neurons.
@@
 
-In the experiment, a perceptron was configured to take complex cells as input.
-It was then trained to distinguish between two shapes shown in \figurename~\ref{fig:7a}.
-The shapes were also used to examine the selectivity of V4 in \cite{pasupathy2001}.
+In the experiment, a perceptron is configured to take complex cells as input.
+It is then trained to distinguish between two shapes shown in \figurename~\ref{fig:7a}.
+The shapes have been used to examine the selectivity of V4 neurons in \cite{pasupathy2001}.
 The difference between the two shapes is that one has a sharp corner towards the top right.
@@
 \caption{Response map and weight matrices of a perceptron that distinguishes two shapes.
-(a) Responses of the perceptron over 18 samples.
-The perceptron prefers the shape in the left half of the samples.
-It is insensitive to stimulus position.
+(a) The perceptron prefers the shape in the left half of the samples.
+It is insensitive to small changes in the position of the shape.
 (b) The input weight of the perceptron. 
@@
 
-Since V4 neurons show a certain degree of invariance in stimulus position,
-we moved the shapes randomly within the receptive field of the perceptron 
+To examine the invariance to small changes in position,
+we move the shapes randomly within the receptive field of the perceptron 
 to generate samples for training and testing (\figurename~\ref{fig:7a} shows several samples).
-The samples were then used as input of the layers of simple cells and complex cells.
-The output of complex cells were passed to the perceptron.
-We had complex cells with 18 different preferred orientations (from $0^\circ$ to $170^\circ$ in steps of $10^\circ$) 
-and over different positions in the V4 receptive field.
-Therefore, the input of the perceptron consisted of 18 matrices, 
+The samples are then used as input of the layers of simple cells and complex cells.
+The output of complex cells are passed to the perceptron.
+Complex cells with 18 different preferred orientations (from $0^\circ$ to $170^\circ$ in steps of $10^\circ$) 
+cover different positions in the V4 receptive field.
+Therefore, the input of the perceptron consists of 18 matrices, 
 each corresponding to the output of complex cells with a certain orientation.
-The input weight of the perceptron was thus also 18 matrices.
+The input weight of the perceptron is also 18 matrices.
 
-The trained perceptron exhibited a strong bias towards the shape with a sharp corner towards the top right.
-It also showed a significant degree of invariance to the stimulus position.
+The trained perceptron exhibits a strong bias towards the shape with a sharp corner towards the top right.
+It also shows a significant degree of invariance to the stimulus position.
 It is obvious that this kind of selectivity is not formed from certain excitatory sub-regions
-or inhibitory sub-regions of the receptive field which was found of simple cells \cite{hubel1962,pasupathy2001}.
+or inhibitory sub-regions of the receptive field which has been found of simple cells \cite{hubel1962,pasupathy2001}.
 The selectivity of the perceptron tallies with the selectivity of actual V4 neurons.
@@
 
-In the following experiment, we trained our model to learn shapes.
-After training, our model can distinguish the stimuli which V4 neurons are selectively tuned for.
-The training samples were collected from \cite{gallant1996,pasupathy1999,pasupathy2001}, 
+In the following experiment, the proposed model is trained to exhibit the shape selectivity found of V4 neurons.
+The training samples are collected from \cite{gallant1996,pasupathy1999,pasupathy2001}, 
 including 4 categories of stimuli, i.e., sinusoidal gratings, non-Cartesian gratings, segmented curves, and closed shapes.
-The training process involved the layers of simple cells, complex cells and RBM encoders.
-Since each sample image was fitted into a single V4 receptive field, the saliency filters were not involved in this experiment.
+The training process involves the layers of simple cells, complex cells and RBM encoders.
+Since each sample image is fitted into a single V4 receptive field, the saliency filters are not involved in this experiment.
 
-We used an RBM with 256 hidden units in this experiment.
-In order to assess the selectivity of these units, we assigned each unit a selectivity index over a category of stimuli.
+An RBM with 256 hidden units is used in this experiment.
+In order to assess the selectivity of these units, we assign each unit a selectivity index over a category of stimuli.
 The selectivity index is defined as the ratio of the maximal response to the average response over the stimuli of a certain category.
@@
 \end{equation}
-In the following statistics, we assumed that a unit has significant selectivity over a category of stimuli
+In the following statistics, it is assumed that a unit has significant selectivity over a category of stimuli
 if its selectivity index is greater than the average index.
-Among the 256 units, 171 units exhibited significant selectivity for different stimuli.
-102 units showed selectivity for segmented curves (\figurename~\ref{fig:8}).
-55 units showed strong bias towards non-Cartesian gratings (\figurename~\ref{fig:9})
-while 50 units showed selectivity for classical sinusoidal gratings.
-68 units exhibited selectivity for closed shapes (\figurename~\ref{fig:10}).
+Among the 256 units, 171 units exhibit significant selectivity for different stimuli.
+102 units show selectivity for segmented curves (\figurename~\ref{fig:8}).
+55 units show strong bias towards non-Cartesian gratings (\figurename~\ref{fig:9})
+while 50 units show selectivity for classical sinusoidal gratings.
+68 units exhibit selectivity for closed shapes (\figurename~\ref{fig:10}).
 The shape selectivity of our model is listed in \tablename~\ref{tab:1}.
@@
 
-\figurename~\ref{fig:8} show three response maps to segmented curves.
+\figurename~\ref{fig:8} shows three response maps to segmented curves.
 Darker background indicates stronger response.
-It is shown that the three units exhibited strong bias towards curves rather than straight lines.
-They were all tuned for the orientation of the projections.
+It is shown that the three units exhibit strong bias towards curves rather than straight lines.
+They are all tuned for the orientation of the projections.
 This is compliant with the selectivity of V4 neurons described in \cite{pasupathy1999}.
 \figurename~\ref{fig:9} shows the response maps of two units which are selective for non-Cartesian gratings.
-Both of the two units were not sensitive to sinusoidal gratings.
+Both of the two units are not sensitive to sinusoidal gratings.
 \figurename~\ref{fig:10} shows the sample patterns of closed shapes and the response map of two units selectively tuned for two kinds of shapes.
-They were sensitive to the angular position of the shapes \cite{pasupathy2001}.
+They are sensitive to the angular position of the shapes \cite{pasupathy2001}.
 
@@
 In this section, we demonstrate the performance of the features generated by our model in feature matching tasks.
-We do not use typical criteria for evaluating local descriptors \cite{mikolajczyk2005}
-for the reason that the typical criteria focus on matching the same physical point in different viewing conditions,
+We do not use typical criteria to evaluate local descriptors \cite{mikolajczyk2005}
+because the typical criteria focus on matching the same physical point in different viewing conditions,
 at which humans or other intelligent species are not significantly better than computer systems.
@@
 The SIFT features \cite{lowe1999} are used for comparison.
-The VLFeat open-source library \cite{vedaldi2010} were utilized for the SIFT implementation and the feature matching algorithm.
+The VLFeat open-source library \cite{vedaldi2010} are utilized for the SIFT implementation and the feature matching algorithm.
 
@@
 more clearly the advantage of our model compared with SIFT feature.
-\figurename~\ref{fig:12a} shows that our model produced correct matches
+\figurename~\ref{fig:12a} shows that our model produces correct matches
 between the model van in the left and the two real vans in the right. 
-However, SIFT feature failed to produce correct matches (\figurename~\ref{fig:12b}). 
-SIFT feature utilizes the direction of image gradient. 
+However, SIFT descriptor fails to produce correct matches (\figurename~\ref{fig:12b}). 
+SIFT descriptor utilizes the direction of image gradient. 
 It fails to capture the shape feature of an object when the object's color changes dramatically because the direction of image gradient changes with the color. 
@@
 They are all cropped to the same size and the same face location.
+A match is correct if it is a pair of corresponding regions in the same position of two images, which contain the same facial feature.
 The images are randomly paired for feature matching.
-A correct match between facial features should be corresponding regions in the two images.
 We use precision-recall curves to compare our neural features with the SIFT features.
@@
 The result is shown in \figurename~\ref{fig:13}.
-Extensive evaluation on the generalization ability of our features is in the object recognition experiment in the next subsection.
+Extensive evaluation on the generalization ability of our features is demonstrated with object recognition experiment in the next subsection.
 
@@
 We evaluate the performance of the proposed model in object recognition in clutter for which 
-the target object in both the training and test samples appears at variable scales and positions within an unsegmented image.
-Features at the salient-points are extracted and passed to a classifier.
-We experimented with the SVM classifier and evaluated the classifier over the Caltech 101 dataset \cite{fei2007}.
+the target object in both the training and test samples appears at variable scales and positions.
+The Caltech 101 dataset \cite{fei2007} is used in this experiment.
+Given an input image, features at the salient-points are extracted and passed to an SVM classifier.
+We evaluate our model with different configurations of parameters.
 
@@
 With a small density, we can lower the dimension of the input of RBM encoders without a significant loss in object recognition performance.
-Thus in the succeeding experiments, a sampling density of $3$ is used.
+Thus in the following experiments, a sampling density of $3$ is used.
 
@@
 The result is shown in \figurename~\ref{fig:14b}.
-With 50 features per image the result approaches the best performance.
+With 50 features per image the performance approaches the upper bound.
 In the succeeding experiments, we take 50 features from each image.
@@
 On the contrary, our proposed model simulates the biological vision in the level of neural mechanism.
-Its multilayer structure resembles the neural systems and extracts information from simple features to gradually more complex features.
+Its multilayer structure resembles the neural systems and gradually extracts information of complex features.
 
@@
 V4 features are automatically extracted based on the output of the artificially designed network.
-This strategy saves us from retraining the first layers every time.
 Additionally, we implement a layer of saliency filters to detect key-points in images,
 which enables the model to process images of various sizes efficiently. 
-In \tablename~\ref{tab:2}, we summarize the comparison of the three models.
+In \tablename~\ref{tab:2}, we summarize the comparison of these three models.
 
-Future work should involve further investigation into the visual systems and extensive evaluation on shape-based object detection.
+Future work will involve further investigation into the visual systems and extensive evaluation on shape-based object detection.
 

### 2015-09-09 article.tex
fix some typo

@@
 Though the neural mechanism of vision has not been fully revealed,
-recent research on neuroscience do provide many inspiring discoveries which would help development of computer vision algorithms.
+recent research on neuroscience provides many inspiring discoveries which would help development of computer vision algorithms.
 In this paper, we demonstrate a novel model of visual area V4 and its application to shape-based feature extraction and object discrimination.
@@
 
-V4 neurons have been examine with different varieties of visual stimuli, including light bars, sinusoidal gratings~(\figurename~\ref{fig:2a}),
+V4 neurons have been examined with different varieties of visual stimuli, including light bars, sinusoidal gratings~(\figurename~\ref{fig:2a}),
 non-Cartesian gratings~(\figurename~\ref{fig:2b}), and closed shapes~(\figurename~\ref{fig:2c}) \cite{gallant1996,pasupathy2001}.
@@
 the kernels of the data-connected layer resemble the pattern of the receptive field of neurons in the primary visual cortex.
-By training many stacked layers of RBM, highly abstracted features can be extracted automatically from input images.
+By training many stacked layers of RBM, highly abstracted features can be automatically extracted from input images.
 Therefore, with additional fine tuning, the deep model exhibits high performance in image classification \cite{krizhevsky2012}.

### 2015-09-10 article.tex
fix image path in latex

@@
 \usepackage[sort&compress]{natbib}
-\usepackage{setspace}
 
@@
 \begin{figure}[!t]
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig1.jpg}} 
+\centerline{\includegraphics[width=0.8\linewidth]{fig1.jpg}} 
 \caption{V4 is an intermediate stage in visual recognition.}
@@
 \centering
-\subfloat[]{\includegraphics[width=0.18\linewidth]{images/fig2a.jpg}\label{fig:2a}}\hfil
-\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig2b.jpg}\label{fig:2b}}\hfil
-\subfloat[]{\includegraphics[width=0.36\linewidth]{images/fig2c.jpg}\label{fig:2c}}
+\subfloat[]{\includegraphics[width=0.18\linewidth]{fig2a.jpg}\label{fig:2a}}\hfil
+\subfloat[]{\includegraphics[width=0.36\linewidth]{fig2b.jpg}\label{fig:2b}}\hfil
+\subfloat[]{\includegraphics[width=0.36\linewidth]{fig2c.jpg}\label{fig:2c}}
 \caption{Shapes to examine V4 selectivity.
@@
 \begin{figure}[!t]
-\centerline{\includegraphics[width=\linewidth]{images/fig3.eps}} 
+\centerline{\includegraphics[width=\linewidth]{fig3.eps}} 
 \caption{Multilayer neural network model of V4.}
@@
 \begin{figure}[!t]
-\centerline{\includegraphics[width=0.8\linewidth]{images/fig4.jpg}} 
+\centerline{\includegraphics[width=0.8\linewidth]{fig4.jpg}} 
 \caption{Entropy of image patches.
@@
 \centering
-\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{images/fig5a.jpg}\label{fig:5a}}\hfil
-\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig5c.jpg}\label{fig:5b}}\hfil
-\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig5e.jpg}\label{fig:5c}}\\
-\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{images/fig5b.jpg}\label{fig:5d}}\hfil
-\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{images/fig5d.jpg}\label{fig:5e}}\hfil
-\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{images/fig5f.jpg}\label{fig:5f}}\\
-\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig5g.jpg}\label{fig:5g}}\hfil
-\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{images/fig5h.jpg}\label{fig:5h}}
+\subfloat[Car image]{\includegraphics[width=0.32\linewidth]{fig5a.jpg}\label{fig:5a}}\hfil
+\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{fig5c.jpg}\label{fig:5b}}\hfil
+\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{fig5e.jpg}\label{fig:5c}}\\
+\subfloat[Face image]{\includegraphics[width=0.32\linewidth]{fig5b.jpg}\label{fig:5d}}\hfil
+\subfloat[Activity map]{\includegraphics[width=0.32\linewidth]{fig5d.jpg}\label{fig:5e}}\hfil
+\subfloat[Entropy map]{\includegraphics[width=0.32\linewidth]{fig5f.jpg}\label{fig:5f}}\\
+\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{fig5g.jpg}\label{fig:5g}}\hfil
+\subfloat[Salient points]{\includegraphics[width=0.49\linewidth]{fig5h.jpg}\label{fig:5h}}
 \caption{Detecting salient points. (a)(d) Original images.
@@
 \begin{figure}[!t]
-\centerline{\includegraphics[width=0.99\linewidth]{images/fig6.eps}} 
+\centerline{\includegraphics[width=0.99\linewidth]{fig6.eps}} 
 \caption{RBM encoder for local shape feature.}
@@
 \centering
-\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{images/fig7a.jpg}\label{fig:7a}}\hfil
-\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{images/fig7b.jpg}\label{fig:7b}}
+\subfloat[Response map]{\includegraphics[width=0.48\linewidth]{fig7a.jpg}\label{fig:7a}}\hfil
+\subfloat[Weight matrices]{\includegraphics[width=0.48\linewidth]{fig7b.jpg}\label{fig:7b}}
 \caption{Response map and weight matrices of a perceptron that distinguishes two shapes.
@@
 \centering
-\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8a.jpg}\label{fig:8a}}\hfil
-\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8b.jpg}\label{fig:8b}}\hfil
-\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8c.jpg}\label{fig:8c}}\hfil
-\subfloat[]{\includegraphics[width=0.24\linewidth]{images/fig8d.jpg}\label{fig:8d}}
+\subfloat[]{\includegraphics[width=0.24\linewidth]{fig8a.jpg}\label{fig:8a}}\hfil
+\subfloat[]{\includegraphics[width=0.24\linewidth]{fig8b.jpg}\label{fig:8b}}\hfil
+\subfloat[]{\includegraphics[width=0.24\linewidth]{fig8c.jpg}\label{fig:8c}}\hfil
+\subfloat[]{\includegraphics[width=0.24\linewidth]{fig8d.jpg}\label{fig:8d}}
 \caption{RBM output units responding to convex curves towards certain directions.
@@
 \subfloat[Response map of a unit selective for a pair of circular sectors (selectivity index = 3.92).]%
-{\includegraphics[width=0.9\linewidth]{images/fig9a.jpg}\label{fig:9a}}\\
+{\includegraphics[width=0.9\linewidth]{fig9a.jpg}\label{fig:9a}}\\
 \subfloat[Response map of a unit selective for helix shapes (Selectivity index = 6.37).]%
-{\includegraphics[width=0.9\linewidth]{images/fig9b.jpg}\label{fig:9b}}
+{\includegraphics[width=0.9\linewidth]{fig9b.jpg}\label{fig:9b}}
 \caption{Response maps of RBM units tuned for non-Cartesian gratings.}
@@
 \centering
-\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{images/fig10a.jpg}\label{fig:10a}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig10b.jpg}\label{fig:10b}}\hfil
-\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{images/fig10c.jpg}\label{fig:10c}}
+\subfloat[Patterns]{\includegraphics[width=0.31\linewidth]{fig10a.jpg}\label{fig:10a}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{fig10b.jpg}\label{fig:10b}}\hfil
+\subfloat[Response map]{\includegraphics[width=0.31\linewidth]{fig10c.jpg}\label{fig:10c}}
 \caption{RBM units responding to closed shapes.}
@@
   \subfloat[Matching between the same face]{
-    \includegraphics[width=0.8\linewidth]{images/fig11a.jpg}
+    \includegraphics[width=0.8\linewidth]{fig11a.jpg}
     \label{fig:11a}}\\
   \subfloat[Matching between different faces]{
-    \includegraphics[width=0.8\linewidth]{images/fig11b.jpg}
+    \includegraphics[width=0.8\linewidth]{fig11b.jpg}
     \label{fig:11b}}
@@
   \subfloat[Our match]{
-    \includegraphics[width=0.8\linewidth]{images/fig12a.jpg}
+    \includegraphics[width=0.8\linewidth]{fig12a.jpg}
     \label{fig:12a}}\\
   \subfloat[SIFT match]{
-    \includegraphics[width=0.8\linewidth]{images/fig12b.jpg}
+    \includegraphics[width=0.8\linewidth]{fig12b.jpg}
     \label{fig:12b}}
@@
 \begin{figure}[!t]
-\centerline{\includegraphics[width=0.9\linewidth]{images/fig13.eps}} 
+\centerline{\includegraphics[width=0.9\linewidth]{fig13.eps}} 
 \caption{Matching facial features in Caltech-101 ``Face-easy'' category.}
@@
   \subfloat[Sampling density of orientations]{
-    \includegraphics[width=0.9\linewidth]{images/fig14a.eps}
+    \includegraphics[width=0.9\linewidth]{fig14a.eps}
     \label{fig:14a}}\\
   \subfloat[Number of features per image]{
-    \includegraphics[width=0.9\linewidth]{images/fig14b.eps}
+    \includegraphics[width=0.9\linewidth]{fig14b.eps}
     \label{fig:14b}}
@@
   \subfloat[Comparison with other methods]{
-    \includegraphics[width=0.9\linewidth]{images/fig15a.eps}
+    \includegraphics[width=0.9\linewidth]{fig15a.eps}
     \label{fig:15a}}\\
   \subfloat[Comparison with other features]{
-    \includegraphics[width=0.9\linewidth]{images/fig15b.eps}
+    \includegraphics[width=0.9\linewidth]{fig15b.eps}
     \label{fig:15b}}
